---
title: "Extended Data Figure 4"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(pheatmap)
library(fgsea)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
```

# Load the data
```{r}
a549_egf <- readRDS("../data/A549_EGF.rds")
a549_tnf <- readRDS("../data/A549_TNF.rds")
a549_tgfb1 <- readRDS("../data/A549_TGFB1.rds")

du145_egf <- readRDS("../data/DU145_EGF.rds")
du145_tnf <- readRDS("../data/DU145_TNF.rds")
du145_tgfb1 <- readRDS("../data/DU145_TGFB1.rds")

mcf7_egf <- readRDS("../data/MCF7_EGF.rds")
mcf7_tnf <- readRDS("../data/MCF7_TNF.rds")
mcf7_tgfb1 <- readRDS("../data/MCF7_TGFB1.rds")

ovca420_egf <- readRDS("../data/OVCA420_EGF.rds")
ovca420_tnf <- readRDS("../data/OVCA420_TNF.rds")
ovca420_tgfb1 <- readRDS("../data/OVCA420_TGFB1.rds")
```


# Extended Data Figure 4a
## Heatmap of hallmarks
```{r}
a549_tgfb1_gam <- read.csv("../output/pseudotime/a549_tgfb1_dge.csv", row.names = 1)
a549_egf_gam <- read.csv("../output/pseudotime/a549_egf_dge.csv", row.names=1)
a549_tnf_gam <- read.csv("../output/pseudotime/a549_tnf_dge.csv", row.names=1)

du145_tgfb1_gam <- read.csv("../output/pseudotime/du145_tgfb1_dge.csv", row.names=1)
du145_egf_gam <- read.csv("../output/pseudotime/du145_egf_dge.csv", row.names=1)
du145_tnf_gam <- read.csv("../output/pseudotime/du145_tnf_dge.csv", row.names=1)

mcf7_tgfb1_gam <- read.csv("../output/pseudotime/mcf7_tgfb1_dge.csv", row.names=1)
mcf7_egf_gam <- read.csv("../output/pseudotime/mcf7_egf_dge.csv", row.names=1)
mcf7_tnf_gam <- read.csv("../output/pseudotime/mcf7_tnf_dge.csv", row.names=1)

ovca420_tgfb1_gam <- read.csv("../output/pseudotime/ovca420_tgfb1_dge.csv", row.names=1)
ovca420_egf_gam <- read.csv("../output/pseudotime/ovca420_egf_dge.csv", row.names=1)
ovca420_tnf_gam <- read.csv("../output/pseudotime/ovca420_tnf_dge.csv", row.names=1)
```

Get sig genes
```{r}
a549_tgfb1_sig <- rownames(a549_tgfb1_gam)[which(a549_tgfb1_gam$fdr <= 0.05)]
a549_egf_sig <- rownames(a549_egf_gam)[which(a549_egf_gam$fdr <= 0.05)]
a549_tnf_sig <- rownames(a549_tnf_gam)[which(a549_tnf_gam$fdr <= 0.05)]

du145_tgfb1_sig <- rownames(du145_tgfb1_gam)[which(du145_tgfb1_gam$fdr <= 0.05)]
du145_egf_sig <- rownames(du145_egf_gam)[which(du145_egf_gam$fdr <= 0.05)]
du145_tnf_sig <- rownames(du145_tnf_gam)[which(du145_tnf_gam$fdr <= 0.05)]

mcf7_tgfb1_sig <- rownames(mcf7_tgfb1_gam)[which(mcf7_tgfb1_gam$fdr <= 0.05)]
mcf7_egf_sig <- rownames(mcf7_egf_gam)[which(mcf7_egf_gam$fdr <= 0.05)]
mcf7_tnf_sig <- rownames(mcf7_tnf_gam)[which(mcf7_tnf_gam$fdr <= 0.05)]

ovca420_tgfb1_sig <- rownames(ovca420_tgfb1_gam)[which(ovca420_tgfb1_gam$fdr <= 0.05)]
ovca420_egf_sig <- rownames(ovca420_egf_gam)[which(ovca420_egf_gam$fdr <= 0.05)]
ovca420_tnf_sig <- rownames(ovca420_tnf_gam)[which(ovca420_tnf_gam$fdr <= 0.05)]
```

Many genes are significant with very modest effect sizes. To try to make it a little more manageable, we'll filter for only genes that are also in the top 2k variable genes

```{r}
#Make a function so we can tweak the number of genes
getVarGenes <- function(seurat_object, n){
  var_table <- seurat_object[["RNA"]]@meta.features
  var_genes <- rownames(var_table[order(desc(var_table$vst.variance.standardized)),])[1:n]
}
```

```{r}
a549_tgfb1_sig <- a549_tgfb1_sig[which(a549_tgfb1_sig %in% getVarGenes(a549_tgfb1, 2000))]
a549_egf_sig <- a549_egf_sig[which(a549_egf_sig %in% getVarGenes(a549_egf, 2000))]
a549_tnf_sig <- a549_tnf_sig[which(a549_tnf_sig %in% getVarGenes(a549_tnf, 2000))]

du145_tgfb1_sig <- du145_tgfb1_sig[which(du145_tgfb1_sig %in% getVarGenes(du145_tgfb1, 2000))]
du145_egf_sig <- du145_egf_sig[which(du145_egf_sig %in% getVarGenes(du145_egf, 2000))]
du145_tnf_sig <- du145_tnf_sig[which(du145_tnf_sig %in% getVarGenes(du145_tnf, 2000))]

mcf7_tgfb1_sig <- mcf7_tgfb1_sig[which(mcf7_tgfb1_sig %in% getVarGenes(mcf7_tgfb1, 2000))]
mcf7_egf_sig <- mcf7_egf_sig[which(mcf7_egf_sig %in% getVarGenes(mcf7_egf, 2000))]
mcf7_tnf_sig <- mcf7_tnf_sig[which(mcf7_tnf_sig %in% getVarGenes(mcf7_tnf, 2000))]

ovca420_tgfb1_sig <- ovca420_tgfb1_sig[which(ovca420_tgfb1_sig %in% getVarGenes(ovca420_tgfb1, 2000))]
ovca420_egf_sig <- ovca420_egf_sig[which(ovca420_egf_sig %in% getVarGenes(ovca420_egf, 2000))]
ovca420_tnf_sig <- ovca420_tnf_sig[which(ovca420_tnf_sig %in% getVarGenes(ovca420_tnf, 2000))]
```

```{r}
dge_list <- data.frame(Gene = unique(c(a549_tgfb1_sig,
                                       a549_egf_sig,
                                       a549_tnf_sig,
                                       du145_tgfb1_sig,
                                       du145_egf_sig,
                                       du145_tnf_sig,
                                       mcf7_tgfb1_sig,
                                       mcf7_egf_sig,
                                       mcf7_tnf_sig,
                                       ovca420_tgfb1_sig,
                                       ovca420_egf_sig,
                                       ovca420_tnf_sig)))
  
dge_list$A549_EGF <- dge_list$Gene %in% a549_egf_sig
dge_list$A549_TGFB1 <- dge_list$Gene %in% a549_tgfb1_sig
dge_list$A549_TNF <- dge_list$Gene %in% a549_tnf_sig

dge_list$DU145_EGF <- dge_list$Gene %in% du145_egf_sig
dge_list$DU145_TGFB1 <- dge_list$Gene %in% du145_tgfb1_sig
dge_list$DU145_TNF <- dge_list$Gene %in% du145_tnf_sig

dge_list$MCF7_EGF <- dge_list$Gene %in% mcf7_egf_sig
dge_list$MCF7_TGFB1 <- dge_list$Gene %in% mcf7_tgfb1_sig
dge_list$MCF7_TNF <- dge_list$Gene %in% mcf7_tnf_sig

dge_list$OVCA420_EGF <- dge_list$Gene %in% ovca420_egf_sig
dge_list$OVCA420_TGFB1 <- dge_list$Gene %in% ovca420_tgfb1_sig
dge_list$OVCA420_TNF <- dge_list$Gene %in% ovca420_tnf_sig

#Convert TRUE/FALSE to 1/0
dge_list[,2:ncol(dge_list)] <- lapply(dge_list[,2:ncol(dge_list)], as.numeric)
```

```{r}
library(fgsea)
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
emt_genes <- hallmarks[["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION"]]
```

```{r}
counts <- dge_list[,2:ncol(dge_list)]
rownames(counts) <- dge_list$Gene
counts <- counts[emt_genes[which(emt_genes %in% rownames(counts))],]
```

```{r}
count_heatmap <- pheatmap(t(counts),
                          color = c("grey95", "red3"),
                          border_color="black",
                          cluster_cols=T,
                          cluster_rows=F,
                          show_rownames=F,
                          show_colnames=T,
                          legend=F,
                          clustering_method="ward.D2",
                          file="../figure_panels/ExtendedDataFig_4/EMT_Hallmarks_DEG_Heatmap.png",
                          width=20, height=4.65)
```


# Extended Data Figure 4b
## GSEA plots
```{r}
conserved_genes <- read.csv("../output/conserved_upregulated_genes.csv", row.names=1)
conserved_genes <- as.character(conserved_genes$x)
```


```{r}
getVars <- function(sample_id){
  seurat_subset <- subset(epithelial, idents=sample_id)
  seurat_subset <- CreateSeuratObject(seurat_subset[["RNA"]]@counts,
                                      min.cells=round(0.001*ncol(seurat_subset)),
                                      meta.data = seurat_subset@meta.data)
  seurat_subset <- NormalizeData(seurat_subset)
  seurat_subset <- FindVariableFeatures(seurat_subset, selection.method = "vst",
                                        nfeatures=1000)
  var_table <- seurat_subset[["RNA"]]@meta.features
  var_table$Gene <- rownames(var_table)
  var_table <- arrange(var_table, desc(vst.variance.standardized))
  var <- var_table$vst.variance.standardized
  names(var) <- var_table$Gene
  gsea_plot <- fgsea::plotEnrichment(conserved_genes, var) +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.line = element_line(size=0.5),
          axis.ticks = element_blank())
  ggsave(gsea_plot, filename = paste0("../figure_panels/ExtendedDataFig_4/", sample_id, "_ConservedGeneSet_GSEA.pdf"),
         width=1.75, height=1.5)
  gsea_plot
}
```

### Lung Tumours
```{r}
seurat <- readRDS("~/Data/2018-lung-cancer/output/lung_cancer.rds")
epithelial <- subset(seurat, idents=1)
Idents(epithelial) <- epithelial$orig.ident
```

```{r}
Sample3b <- getVars("Sample3b")
Sample3c <- getVars("Sample3c")
Sample4a <- getVars("Sample4a")
Sample4b <- getVars("Sample4b")
Sample4c <- getVars("Sample4c")
Sample4d <- getVars("Sample4d")
Sample5a <- getVars("Sample5a")
Sample5b <- getVars("Sample5b")
Sample5d <- getVars("Sample5d")
```

### Syngeneic tumours
```{r}
getVars <- function(sample_id){
  seurat_subset <- subset(seurat, idents=sample_id)
  seurat_subset <- CreateSeuratObject(seurat_subset[["RNA"]]@counts,
                                      min.cells=round(0.001*ncol(seurat_subset)),
                                      meta.data = seurat_subset@meta.data)
  seurat_subset <- NormalizeData(seurat_subset)
  seurat_subset <- FindVariableFeatures(seurat_subset, selection.method = "vst",
                                        nfeatures=1000)
  var_table <- seurat_subset[["RNA"]]@meta.features
  var_table$Gene <- toupper(rownames(var_table))
  var_table <- arrange(var_table, desc(vst.variance.standardized))
  var <- var_table$vst.variance.standardized
  names(var) <- var_table$Gene
  gsea_plot <- fgsea::plotEnrichment(conserved_genes, var) +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.line = element_line(size=0.5),
          axis.ticks = element_blank())
  ggsave(gsea_plot, filename = paste0("../figure_panels/ExtendedDataFig_4/", sample_id, "_ConservedGeneSet_GSEA.pdf"),
         width=1.75, height=1.5)
  gsea_plot
}
```

```{r}
syngeneic_seurat <- readRDS("~/Data/2019-syngeneic-tumours/output/syngeneic_tumors.rds")
syngeneic_subset <- subset(syngeneic_seurat, idents = c(0,2))
Idents(syngeneic_subset) <- syngeneic_subset$model
seurat <- syngeneic_subset
```

```{r}
B16 <- getVars("B16")
CT26 <- getVars("CT26")
EMT6 <- getVars("EMT6")
LL2 <- getVars("LL2")
MC38 <- getVars("MC38")
SA1N <- getVars("SA1N")
```

### Tabula Muris Data
```{r}
seurat <- readRDS("~/Projects/tabula_muris_analysis/output/tabula_muris.rds")
```

```{r}
Idents(seurat) <- seurat$cell_ontology_class
```

```{r}
bladder_epithelium <- getVars("bladder cell")
bladder_basal <- getVars("basal cell of urothelium")
lung_mesothelial <- getVars("mesothelial cell")
mammary_basal <- getVars("basal cell")
tongue_basal <- getVars("basal cell of epidermis")
trachea_epithelial <- getVars("epithelial cell")
trachea_basal <- getVars("basal cell of epithelium of trachea")
```

# Extended Data Fig 4c
## Correlation of conserved genes
```{r}
getCor <- function(seurat_object){
  exp <- as.matrix(seurat_object[["RNA"]]@data)
  rownames(exp) <- toupper(rownames(exp))
  genes_keep <- up_genes[which(up_genes %in% rownames(exp))]
  score_cor <- mean(cor(exp[genes_keep,], method="spearman"))
}
```

Get seurat objects for datasets
Tumors
```{r}
Sample3b <- subset(epithelial, idents = "Sample3b")
Sample3c <- subset(epithelial, idents = "Sample3c")
Sample4a <- subset(epithelial, idents = "Sample4a")
Sample4b <- subset(epithelial, idents = "Sample4b")
Sample4c <- subset(epithelial, idents = "Sample4c")
Sample5a <- subset(epithelial, idents = "Sample5a")
Sample5b <- subset(epithelial, idents = "Sample5b")
Sample5d <- subset(epithelial, idents = "Sample5d")
```

Syngeneic Models
```{r}
B16 <- subset(syngeneic_subset, idents = "B16")
CT26 <- subset(syngeneic_subset, idents = "CT26")
EMT6 <- subset(syngeneic_subset, idents = "EMT6")
LL2 <- subset(syngeneic_subset, idents = "LL2")
MC38 <- subset(syngeneic_subset, idents = "MC38")
SA1N <- subset(syngeneic_subset, idents = "SA1N")
```

Tabula
```{r}
bladder_epithelium <- subset(seurat, idents="bladder cell")
bladder_basal <- subset(seurat, idents="basal cell of urothelium")
lung_mesothelial <- subset(seurat, idents="mesothelial cell")
mammary_basal <- subset(seurat, idents="basal cell")
tongue_basal <- subset(seurat, idents="basal cell of epidermis")
trachea_epithelial <- subset(seurat, idents="epithelial cell")
trachea_basal <- subset(seurat, idents="basal cell of epithelium of trachea")
```

```{r}
seurat_list <- list(a549_tgfb1, a549_egf, a549_tnf, du145_tgfb1, du145_egf, du145_tnf,
                 mcf7_tgfb1, mcf7_egf, mcf7_tnf, ovca420_tgfb1, ovca420_egf, ovca420_tnf,
                 Sample3b, Sample3c, Sample4a, Sample4b, Sample4c, Sample5a, Sample5b, Sample5d,
                 B16, CT26, EMT6, LL2, MC38, SA1N, bladder_epithelium, bladder_basal,
                 lung_mesothelial, mammary_basal, tongue_basal, trachea_epithelial,
                 trachea_basal)
```

```{r}
correlations <- lapply(seurat_list, getCor)
correlations <- do.call(rbind, correlations)
correlations <- as.matrix(correlations)
rownames(correlations) <- c("EMT_A549_TGFB1", "EMT_A549_EGF", "EMT_A549_TNF",
                            "EMT_DU145_TGFB1", "EMT_DU145_EGF", "EMT_DU145_TNF",
                            "EMT_MCF7_TGFB1", "EMT_MCF7_EGF", "EMT_MCF7_TNF",
                            "EMT_OVCA420_TGFB1", "EMT_OVCA420_EGF", "EMT_OVCA420_TNF",
                            "LungTumour_3b", "LungTumour_3c", "LungTumour_4a", "LungTumour_4b",
                            "LungTumour_4c", "LungTumour_5a", "LungTumour_5b", "LungTumour_5d",
                            "MouseTumour_B16", "MouseTumour_CT26", "MouseTumour_EMT6",
                            "MouseTumour_LL2", "MouseTumour_MC38", "MouseTumour_SA1N",
                            "MouseTissue_BladderEpithelium", "MouseTissue_BladderBasal", 
                            "MouseTissue_LungMesothelial", "MouseTissue_MammaryBasal",
                            "MouseTissue_TongueBasal", "MouseTissue_TracheaEpithelium",
                            "MouseTissue_TracheaBasal")
```

```{r}
color_hack <- round((min(correlations) * 100) +100) #makes it so that the RdBu palette goes from -1 to 1, but we only plot the relevant range 
correlation_heatmap <- pheatmap(correlations,
                                color = colorRampPalette(rev(brewer.pal(7, "RdBu")))(200)[color_hack:200],
                                legend=F,
                                border_color="black",
                                cluster_cols=F,
                                cluster_rows=F,
                                show_rownames=T,
                                show_colnames=F,
                                display_numbers = T,
                                number_color = "grey15",
                                filename = "../figure_panels/ExtendedDataFig_4/Heatmap_EMT_UP_correlation.png",
                                width=2.65, height=8.25)
```

