---
title: "Timecourse - Subsetting data"
output: html_notebook
---

# Goal

# Dependencies
```{r}
library(Seurat)
library(dplyr)
library(tidyr)
library(viridis)
library(matrixStats)
```

# Load data
```{r}
seurat <- readRDS("../data/seurat_unfiltered.rds")
cc.genes <- readLines(con = "~/Data/GeneLists/regev_lab_cell_cycle_genes.txt")
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
```


# Explore full data
Before we subset the data, we'll do a quick look at all of the data and cluster it
to help subset it

```{r}
Dark2 <- RColorBrewer::brewer.pal(8, "Dark2")
Set1 <- RColorBrewer::brewer.pal(8, "Set1")
Set2 <- RColorBrewer::brewer.pal(8, "Set2")
```


## Counts for each sample
```{r}
data <- seurat@meta.data
data <- group_by(data, CellLine, Treatment) %>%
  summarise(Count=n())
data <- na.omit(data)
data$Treatment <- factor(data$Treatment, levels=c("TGFB1", "EGF", "TNF"))

data_plot <- ggplot(data, aes(x=Treatment, y=Count)) +
  geom_col(color="black", aes(fill=CellLine)) +
  facet_wrap(~CellLine, scales="free") +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_manual(values=Dark2[1:4]) +
  ylab("Cell Count") + xlab("") +
  theme_classic() +
  coord_flip() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size=12),
        legend.position="none",
        axis.text=element_text(size=10, color="black"),
        axis.title=element_text(size=12))
ggsave(data_plot, filename = "../figs/Timecourse_CellCounts.png",
       width=5.5, height=4.25)
```


```{r}
plot_data <- seurat@meta.data
plot_data$UMAP1 <- Embeddings(seurat, reduction="umap")[,1]
plot_data$UMAP2 <- Embeddings(seurat, reduction="umap")[,2]
plot_data$Doublet <- factor(plot_data$Doublet,
                            levels=c("Singlet", "Doublet", "Negative"))
plot_data$CellLine[is.na(plot_data$CellLine)] <- "NA"
plot_data$CellLine <- factor(plot_data$CellLine, 
                             levels=c("A549", "DU145", "MCF7", "OVCA420", "NA"))
#For facet plot
plot_data_bg <- plot_data
plot_data_bg$CellLine <- NULL

bulk_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.2, alpha=0.8, aes(color=Doublet)) +
  scale_color_manual(values=c("grey", "red", "darkgreen")) +
  theme_void()
ggsave(bulk_plot, file="../figs/MULTIseq_classifications.pdf",
       width=4.5, height=3, useDingbats=F)
ggsave(bulk_plot, file="../figs/MULTIseq_classifications.png",
       width=4.5, height=3)

bulk_sample_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.2, alpha=0.8, aes(color=CellLine)) +
  scale_color_manual(values=c(Dark2[1], Dark2[2], Dark2[3], Dark2[4],  "grey")) +
  theme_void()
ggsave(bulk_sample_plot, file="../figs/MULTIseq_classifications_CellLine.pdf",
       width=4.5, height=3, useDingbats=F)
ggsave(bulk_sample_plot, file="../figs/MULTIseq_classifications_CellLine.png",
       width=4.5, height=3)

bulk_facet_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(data=plot_data_bg, size=0.1, alpha=0.4, color='lightgrey') +
  geom_point(size=0.2, alpha=0.4, aes(color=CellLine)) +
  scale_color_manual(values=c(Dark2[1], Dark2[2], Dark2[3], Dark2[4], Dark2[5], "grey20")) +
  facet_wrap(~CellLine, ncol=5) +
  theme_void() +
  theme(legend.position="none",
          axis.text = element_blank(),
          axis.title = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          strip.background = element_blank())
ggsave(bulk_facet_plot, file="../figs/MULTIseq_classifications_CellLine_facet.pdf",
       width=10, height=2.5, useDingbats=F)
ggsave(bulk_facet_plot, file="../figs/MULTIseq_classifications_CellLine_facet.png",
       width=10, height=2.5)

bulk_plot
bulk_sample_plot
bulk_facet_plot
```

## Frequency of cell annotations in wrong cluster
```{r}
plot_data$Cluster <- Idents(seurat)
bulk_sample_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.1, alpha=0.8, aes(color=Cluster)) +
  scale_colour_manual(values=c(Set1[1], Set1[2], Set1[3], Set1[4], Set1[5], Set1[7])) +
  theme_classic() +
  theme(axis.text=element_blank(),
        axis.line=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank())
ggsave(bulk_sample_plot, file="../figs/UMAP_clustered.pdf",
       width=4.5, height=3, useDingbats=F)
ggsave(bulk_sample_plot, file="../figs/UMAP_clustered.png",
       width=4.5, height=3)
bulk_sample_plot
```

```{r}
seurat$Cluster <- Idents(seurat)
```


```{r}
clusters_freq <- plot_data %>%
  group_by(CellLine, Cluster) %>%
  summarise(n=n()) %>%
  mutate(relative_freq = n/sum(n))
clusters_freq$Cluster <- factor(clusters_freq$Cluster)
clusters_freq <- dplyr::filter(clusters_freq, CellLine != "NA")
clusters_freq$CellLine <- factor(clusters_freq$CellLine,
                                 levels=c("A549", "DU145", "MCF7", "OVCA420"))

freq_plot <- ggplot(clusters_freq, aes(x=CellLine, y=relative_freq)) +
  geom_col(aes(fill=Cluster), position="stack") +
  scale_fill_manual(values=c(Set1[1], Set1[2], Set1[3], Set1[4], Set1[5], Set1[7])) +
  scale_y_continuous(limits=c(0,1), expand=c(0,0)) +
  ylab("Relative Frequency") +
  theme_classic() +
  coord_flip() +
  theme(strip.background=element_blank(),
        axis.line.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size=12, colour="black"),
        axis.text = element_text(size=10, colour="black"),
        #legend.key.size = unit(4, "mm"),
        legend.position="none")
ggsave(freq_plot, file="../figs/Cluster_relative_frequency.pdf",
       width=4, height=1.5)
ggsave(freq_plot, file="../figs/Cluster_relative_frequency.png",
       width=4, height=1.5)
freq_plot
```

# Function to subset data
Note: OVCA420s contain two legitimate clusters. Cluster 4 appears to be a divergent population
based on a more in-depth analysis. This population is unresponsive to any treatments and is not present
in our deeper dataset from older cell stocks. I'll eliminate it largely because it is unresponsive to any treatment


```{r}
dominant_clusters <- data.frame(CellLine = c("A549", "DU145", "MCF7", "OVCA420"),
                                Cluster = c(0, 1, 2, 3))
```

```{r}
DefaultAssay(seurat) <- "RNA" #Just to ensure everything following is on RNA assay and not SCT
```

```{r}
getSeuratSubset <- function(cell_line, treatment){
  print(paste("Subsetting data to get",cell_line,"cells, treated with", treatment, "..."))
  cell_filter <- colnames(seurat)[which(seurat$CellLine == cell_line &
                                          (as.numeric(seurat$Cluster)-1) %in% 
                                          dplyr::filter(dominant_clusters, 
                                                        CellLine==cell_line)$Cluster)]
  #That hacky first part of the last line is just because the as.numeric converts 0-4 to 1-5
  genes_filter <- rownames(seurat[["RNA"]]@data)[rowSums(as.matrix(seurat[["RNA"]]@data)[, cell_filter]) > 0]
  
  print("Subsetting data by cell line")
  subset_seurat <- subset(seurat, cells=cell_filter, features=genes_filter)
  
  print("Re-classifying cell cycle") #likely cleanest on pure cell line
  subset_seurat <- CellCycleScoring(subset_seurat, s.features = s.genes, 
                                    g2m.features = g2m.genes, set.ident = FALSE)
  
  print("Subsetting data by treatment")
  cell_filter <- colnames(subset_seurat)[which(subset_seurat$Treatment == treatment)]
  genes_filter <- rownames(subset_seurat[["RNA"]]@data)[rowSums(as.matrix(subset_seurat[["RNA"]]@data)[, cell_filter]) > 0]
  subset_seurat <- subset(subset_seurat, cells=cell_filter, features=genes_filter)
  
  
  print("Normalizing the data")
  subset_seurat <- SCTransform(subset_seurat,
                               do.correct.umi=T,
                               vars.to.regress=c("S.Score", "G2M.Score",
                                                 "percent.mito"),
                               batch_var="Mix",
                               return_gene_attr=T,
                               variable.features.n=2000)
  
  print("Re-leveling sample factor")
  subset_seurat$Sample <- factor(subset_seurat$Sample,
                                     levels=c(paste0(cell_line, "_", treatment, "_0d"),
                                              paste0(cell_line, "_", treatment, "_8h"),
                                              paste0(cell_line, "_", treatment, "_1d"),
                                              paste0(cell_line, "_", treatment, "_3d"),
                                              paste0(cell_line, "_", treatment, "_7d"),
                                              paste0(cell_line, "_", treatment, "_8h_rm"),
                                              paste0(cell_line, "_", treatment, "_1d_rm"),
                                              paste0(cell_line, "_", treatment, "_3d_rm")))
  
  print("Running PCA...")
  subset_seurat <- RunPCA(subset_seurat, verbose=F)
  
  print("Clustering...")
  subset_seurat <- FindNeighbors(subset_seurat, dims=1:30)
  subset_seurat <- FindClusters(subset_seurat, resolution=0.5)
  
  print("Running UMAP...")
  subset_seurat <- RunUMAP(subset_seurat, dims=1:30)
}
```

# Subset the data
```{r}                                                                       
a549_tgfb1 <- getSeuratSubset("A549", "TGFB1")
a549_tnf <- getSeuratSubset("A549", "TNF") 
a549_egf <- getSeuratSubset("A549", "EGF") 

du145_tgfb1 <- getSeuratSubset("DU145", "TGFB1")
du145_tnf <- getSeuratSubset("DU145", "TNF") 
du145_egf <- getSeuratSubset("DU145", "EGF")

mcf7_tgfb1 <- getSeuratSubset("MCF7", "TGFB1")
mcf7_tnf <- getSeuratSubset("MCF7", "TNF")
mcf7_egf <- getSeuratSubset("MCF7", "EGF")

ovca420_tgfb1 <- getSeuratSubset("OVCA420", "TGFB1")
ovca420_tnf <- getSeuratSubset("OVCA420", "TNF")
ovca420_egf <- getSeuratSubset("OVCA420", "EGF")
```

# Visualize
```{r}
DimPlot(a549_tgfb1, reduction='umap', group.by="Mix")
DimPlot(a549_tgfb1, reduction='umap', group.by='Sample')
DimPlot(a549_tgfb1, reduction='umap')
DimPlot(a549_egf, reduction='umap', group.by="Mix")
DimPlot(a549_egf, reduction='umap', group.by='Sample')
DimPlot(a549_egf, reduction='umap')
DimPlot(a549_tnf, reduction='umap', group.by="Mix")
DimPlot(a549_tnf, reduction='umap', group.by='Sample')
DimPlot(a549_tnf, reduction='umap')
```

```{r}
DimPlot(du145_tgfb1, reduction='umap', group.by="Mix")
DimPlot(du145_tgfb1, reduction='umap', group.by='Sample')
DimPlot(du145_tgfb1, reduction='umap')
DimPlot(du145_egf, reduction='umap',group.by="Mix")
DimPlot(du145_egf, reduction='umap', group.by='Sample')
DimPlot(du145_egf, reduction='umap')
DimPlot(du145_tnf, reduction='umap', group.by="Mix")
DimPlot(du145_tnf, reduction='umap', group.by='Sample')
DimPlot(du145_tnf, reduction='umap')
```

```{r}
DimPlot(mcf7_tgfb1, reduction='umap', group.by="Mix")
DimPlot(mcf7_tgfb1, reduction='umap', group.by='Sample')
DimPlot(mcf7_tgfb1, reduction='umap')
DimPlot(mcf7_egf, reduction='umap', group.by="Mix")
DimPlot(mcf7_egf, reduction='umap', group.by='Sample')
DimPlot(mcf7_egf, reduction='umap')
DimPlot(mcf7_tnf, reduction='umap', group.by="Mix")
DimPlot(mcf7_tnf, reduction='umap', group.by='Sample')
DimPlot(mcf7_tnf, reduction='umap')
```

```{r}
DimPlot(ovca420_tgfb1, reduction='umap', group.by="Mix")
DimPlot(ovca420_tgfb1, reduction='umap', group.by='Sample')
DimPlot(ovca420_tgfb1, reduction='umap')
DimPlot(ovca420_egf, reduction='umap', group.by="Mix")
DimPlot(ovca420_egf, reduction='umap', group.by='Sample')
DimPlot(ovca420_egf, reduction='umap')
DimPlot(ovca420_tnf, reduction='umap', group.by="Mix")
DimPlot(ovca420_tnf, reduction='umap', group.by='Sample')
DimPlot(ovca420_tnf, reduction='umap')
```

# Save subsets
```{r}
saveRDS(a549_tgfb1, file="../data/A549_TGFB1.rds")
saveRDS(a549_egf, file="../data/A549_EGF.rds")
saveRDS(a549_tnf, file="../data/A549_TNF.rds")

saveRDS(du145_tgfb1, file="../data/DU145_TGFB1.rds")
saveRDS(du145_egf, file="../data/DU145_EGF.rds")
saveRDS(du145_tnf, file="../data/DU145_TNF.rds")

saveRDS(ovca420_tgfb1, file="../data/OVCA420_TGFB1.rds")
saveRDS(ovca420_egf, file="../data/OVCA420_EGF.rds")
saveRDS(ovca420_tnf, file="../data/OVCA420_TNF.rds")

saveRDS(mcf7_tgfb1, file="../data/MCF7_TGFB1.rds")
saveRDS(mcf7_egf, file="../data/MCF7_EGF.rds")
saveRDS(mcf7_tnf, file="../data/MCF7_TNF.rds")
```


# Clean up contaminating cells from subsets
```{r}
removeClusters <- function(seurat_object, clusters_remove){
  print("Subsetting data")
  clusters_keep <- unique(as.character(Idents(seurat_object)))
  clusters_keep <- clusters_keep[-which(clusters_keep %in% clusters_remove)]
  seurat_subset <- subset(seurat_object, idents=clusters_keep)
  
  #Prep to re-normalize with SCTransform, without trash clusters
  DefaultAssay(seurat_subset) <- "RNA"
  
  print("Normalizing the data")
  seurat_subset <- SCTransform(seurat_subset,
                               do.correct.umi=T,
                               vars.to.regress=c("S.Score", "G2M.Score",
                                                 "percent.mito"),
                               batch_var="Mix",
                               return_gene_attr=T,
                               variable.features.n=2000)
  
  print("Running PCA...")
  seurat_subset <- RunPCA(seurat_subset, verbose=F)
  
  print("Clustering...")
  seurat_subset <- FindNeighbors(seurat_subset, dims=1:30)
  seurat_subset <- FindClusters(seurat_subset, resolution=0.5)
  
  print("Running UMAP...")
  seurat_subset <- RunUMAP(seurat_subset, dims=1:30)
}
```

```{r}
a549_tgfb1 <- removeClusters(a549_tgfb1, 4)
a549_egf <- removeClusters(a549_egf, c(5,6,8))
a549_tnf <- removeClusters(a549_tnf, c(6,7,9,10,11))

du145_tgfb1 <- removeClusters(du145_tgfb1, c(6,7))
du145_egf <- removeClusters(du145_egf, c(5,6,8))
du145_tnf <- removeClusters(du145_tnf, c(3,4,8))

mcf7_tgfb1 <- removeClusters(mcf7_tgfb1, c(4))
mcf7_egf <- removeClusters(mcf7_egf, c(6,7))
mcf7_tnf <- removeClusters(mcf7_tnf, c(4))

ovca420_egf <- removeClusters(ovca420_egf, c(3,4))
ovca420_tnf <- removeClusters(ovca420_tnf, c(4,5,6))
```

# Re-visualize
```{r}
DimPlot(a549_tgfb1, reduction='umap', group.by="Mix")
DimPlot(a549_tgfb1, reduction='umap', group.by='Sample')
DimPlot(a549_tgfb1, reduction='umap')
DimPlot(a549_egf, reduction='umap', group.by="Mix")
DimPlot(a549_egf, reduction='umap', group.by='Sample')
DimPlot(a549_egf, reduction='umap')
DimPlot(a549_tnf, reduction='umap', group.by="Mix")
DimPlot(a549_tnf, reduction='umap', group.by='Sample')
DimPlot(a549_tnf, reduction='umap')
```

```{r}
DimPlot(du145_tgfb1, reduction='umap', group.by="Mix")
DimPlot(du145_tgfb1, reduction='umap', group.by='Sample')
DimPlot(du145_tgfb1, reduction='umap')
DimPlot(du145_egf, reduction='umap',group.by="Mix")
DimPlot(du145_egf, reduction='umap', group.by='Sample')
DimPlot(du145_egf, reduction='umap')
DimPlot(du145_tnf, reduction='umap', group.by="Mix")
DimPlot(du145_tnf, reduction='umap', group.by='Sample')
DimPlot(du145_tnf, reduction='umap')
```

```{r}
DimPlot(mcf7_tgfb1, reduction='umap', group.by="Mix")
DimPlot(mcf7_tgfb1, reduction='umap', group.by='Sample')
DimPlot(mcf7_tgfb1, reduction='umap')
DimPlot(mcf7_egf, reduction='umap', group.by="Mix")
DimPlot(mcf7_egf, reduction='umap', group.by='Sample')
DimPlot(mcf7_egf, reduction='umap')
DimPlot(mcf7_tnf, reduction='umap', group.by="Mix")
DimPlot(mcf7_tnf, reduction='umap', group.by='Sample')
DimPlot(mcf7_tnf, reduction='umap')
```

```{r}
DimPlot(ovca420_tgfb1, reduction='umap', group.by="Mix")
DimPlot(ovca420_tgfb1, reduction='umap', group.by='Sample')
DimPlot(ovca420_tgfb1, reduction='umap')
DimPlot(ovca420_egf, reduction='umap', group.by="Mix")
DimPlot(ovca420_egf, reduction='umap', group.by='Sample')
DimPlot(ovca420_egf, reduction='umap')
DimPlot(ovca420_tnf, reduction='umap', group.by="Mix")
DimPlot(ovca420_tnf, reduction='umap', group.by='Sample')
DimPlot(ovca420_tnf, reduction='umap')
```

# Subsetting cell lines with aggregated treatments
```{r}
getSeuratSubset <- function(cell_line){
  print(paste("Subsetting data to get",cell_line,"cell"))
  cell_filter <- colnames(seurat)[which(seurat$CellLine == cell_line &
                                          (as.numeric(seurat$Cluster)-1) %in% 
                                          dplyr::filter(dominant_clusters, 
                                                        CellLine==cell_line)$Cluster)]
  #That hacky first part of the last line is just because the as.numeric converts 0-4 to 1-5
  genes_filter <- rownames(seurat[["RNA"]]@data)[rowSums(as.matrix(seurat[["RNA"]]@data)[, cell_filter]) > 0]
  
  print("Subsetting data by cell line")
  subset_seurat <- subset(seurat, cells=cell_filter, features=genes_filter)
  
  print("Re-classifying cell cycle") #likely cleanest on pure cell line
  subset_seurat <- CellCycleScoring(subset_seurat, s.features = s.genes, 
                                    g2m.features = g2m.genes, set.ident = FALSE)
  
  
  print("Normalizing the data")
  subset_seurat <- SCTransform(subset_seurat,
                               do.correct.umi=T,
                               vars.to.regress=c("S.Score", "G2M.Score",
                                                 "percent.mito"),
                               batch_var="Mix",
                               return_gene_attr=T,
                               variable.features.n=5000)
  
  print("Running PCA...")
  subset_seurat <- RunPCA(subset_seurat, verbose=F)
  
  print("Clustering...")
  subset_seurat <- FindNeighbors(subset_seurat, dims=1:30)
  subset_seurat <- FindClusters(subset_seurat, resolution=0.5)
  
  print("Running UMAP...")
  subset_seurat <- RunUMAP(subset_seurat, dims=1:30)
}
```

```{r}
a549 <- getSeuratSubset("A549")
du145 <- getSeuratSubset("DU145")
mcf7 <- getSeuratSubset("MCF7")
ovca420 <- getSeuratSubset("OVCA420")
```

## Visualize embeddings
```{r}
DimPlot(a549, group.by="Treatment")
DimPlot(du145, group.by="Treatment")
DimPlot(mcf7, group.by="Treatment")
DimPlot(ovca420, group.by="Treatment")
```


# Save final subsets

```{r}
saveRDS(a549_tgfb1, file="../data/A549_TGFB1.rds")
saveRDS(a549_egf, file="../data/A549_EGF.rds")
saveRDS(a549_tnf, file="../data/A549_TNF.rds")

saveRDS(du145_tgfb1, file="../data/DU145_TGFB1.rds")
saveRDS(du145_egf, file="../data/DU145_EGF.rds")
saveRDS(du145_tnf, file="../data/DU145_TNF.rds")

saveRDS(ovca420_tgfb1, file="../data/OVCA420_TGFB1.rds")
saveRDS(ovca420_egf, file="../data/OVCA420_EGF.rds")
saveRDS(ovca420_tnf, file="../data/OVCA420_TNF.rds")

saveRDS(mcf7_tgfb1, file="../data/MCF7_TGFB1.rds")
saveRDS(mcf7_egf, file="../data/MCF7_EGF.rds")
saveRDS(mcf7_tnf, file="../data/MCF7_TNF.rds")
```


