---
title: "Processing: Timecourse data"
output: html_notebook
---

# Goal
The goal of this notebook is to 1) take the cellranger output for these time course samples and get them into Seurat objects, and 2) perform LMO counting for each sample and use that to annotate the cells in the seurat object

# Dependencies
```{r}
library(Seurat) #v3.x
library(viridis)
library(dplyr)
library(cowplot)
```

# Load in the Cellranger output
```{r}
mix1_mat <- Read10X(data.dir = "../data/cellranger/Mix1/filtered_gene_bc_matrices/GRCh38/")
mix2_mat <- Read10X(data.dir = "../data/cellranger/Mix2/filtered_gene_bc_matrices/GRCh38/")
mix3a_mat <- Read10X(data.dir="../data/cellranger/Mix3a/filtered_gene_bc_matrices/GRCh38/")
mix3b_mat <- Read10X(data.dir="../data/cellranger/Mix3b/filtered_gene_bc_matrices/GRCh38/")
mix4a_mat <- Read10X(data.dir="../data/cellranger/Mix4a/filtered_gene_bc_matrices/GRCh38/")
mix4b_mat <- Read10X(data.dir="../data/cellranger/Mix4b/filtered_gene_bc_matrices/GRCh38/")
```

```{r}
mix1.cells <- data.frame(cell = colnames(mix1_mat))
write.table(mix1.cells, "../data/Mix1_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)

mix2.cells <- data.frame(cell = colnames(mix2_mat))
write.table(mix2.cells, "../data/Mix2_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)

mix3a.cells <- data.frame(cell = colnames(mix3a_mat))
write.table(mix3a.cells, "../data/Mix3a_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)

mix3b.cells <- data.frame(cell = colnames(mix3b_mat))
write.table(mix3b.cells, "../data/Mix3b_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)

mix4a.cells <- data.frame(cell = colnames(mix4a_mat))
write.table(mix4a.cells, "../data/Mix4a_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)

mix4b.cells <- data.frame(cell = colnames(mix4b_mat))
write.table(mix4b.cells, "../data/Mix4b_cellIDs.txt", sep="\t", col.names=F, 
            row.names=F, quote=F)
```

# Mix 1 QC and basic Processing
This first round of processing is only just to get the data organized. Each condition will be filtered out and reprocessed later

```{r}
seurat <- CreateSeuratObject(mix1_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix1")
```

```{r}
mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.09)
mix1_keep <- colnames(seurat)
mix1_seurat <- seurat
```

# Mix2 QC and basic processing
```{r}
seurat <- CreateSeuratObject(mix2_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix2")

mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.1)
mix2_keep <- colnames(seurat)
mix2_seurat <- seurat
```

# Mix3a QC and basic processing
```{r}
seurat <- CreateSeuratObject(mix3a_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix3a")

mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.1)
mix3a_keep <- colnames(seurat)
mix3a_seurat <- seurat
```

# Mix3b QC and basic processing

```{r}
seurat <- CreateSeuratObject(mix3b_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix3b")

mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.1)
mix3b_keep <- colnames(seurat)
mix3b_seurat <- seurat
```

# Mix4a QC and basic processing

```{r}
seurat <- CreateSeuratObject(mix4a_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix4a")

mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.11)
mix4a_keep <- colnames(seurat)
mix4a_seurat <- seurat
```

# Mix4b QC and basic processing

```{r}
seurat <- CreateSeuratObject(mix4b_mat, min.cells = 3, 
                             min.features = 200,
                             project = "Mix4b")

mito.features <- grep(pattern = "^MT-", x = rownames(seurat), value = TRUE)
percent.mito <- Matrix::colSums(GetAssayData(seurat, slot = 'counts')[mito.features, ]) / Matrix::colSums(GetAssayData(seurat, slot = 'counts'))

# The [[ operator can add columns to object metadata, and is a great place to stash QC stats
seurat[['percent.mito']] <- percent.mito
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mito"), ncol = 3)
```

```{r}
seurat <- subset(seurat, 
                 subset = nFeature_RNA > 200 & 
                   percent.mito < 0.1)
mix4b_keep <- colnames(seurat)
mix4b_seurat <- seurat
```

# Combine data
```{r}
length(c(mix1_keep, mix2_keep, mix3a_keep, mix3b_keep, mix4a_keep, mix4b_keep))
```

```{r}
seurat <- merge(mix1_seurat, list(mix2_seurat, mix3a_seurat,
                             mix3b_seurat, mix4a_seurat, mix4b_seurat),
                add.cell.ids = c("Mix1", "Mix2", "Mix3a", "Mix3b",
                                 "Mix4a", "Mix4b"),
                project = "EMT_Timecourse")
```

```{r}
rm(mix1_seurat, mix2_seurat, mix3a_seurat, mix3b_seurat,
   mix4a_seurat, mix4b_seurat)
```


# Annotate cells based on LMOs
```{r}
bar.ref <- read.csv("../data/LMOlist.csv", stringsAsFactors = F,
                    header=F)$V1
```

## Quantify barcodes from fastq files
```{r}
readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix1-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix1-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix1_keep,
                                 chemistry="V2")
mix1_bar_table <- MULTIseq.align(readTable, mix1_keep, bar.ref)
write.csv(mix1_bar_table, file="../data/Mix1_barcode_counts.csv", quote=F)

readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix2-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix2-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix2_keep,
                                 chemistry="V2")
mix2_bar_table <- MULTIseq.align(readTable, mix2_keep, bar.ref)
write.csv(mix2_bar_table, file="../data/Mix2_barcode_counts.csv", quote=F)

readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix3a-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix3a-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix3a_keep,
                                 chemistry="V2")
mix3a_bar_table <- MULTIseq.align(readTable, mix3a_keep, bar.ref)
write.csv(mix3a_bar_table, file="../data/Mix3a_barcode_counts.csv", quote=F)

readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix3b-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix3b-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix3b_keep,
                                 chemistry="V2")
mix3b_bar_table <- MULTIseq.align(readTable, mix3b_keep, bar.ref)
write.csv(mix3b_bar_table, file="../data/Mix3b_barcode_counts.csv", quote=F)

readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix4a-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix4a-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix4a_keep,
                                 chemistry="V2")
mix4a_bar_table <- MULTIseq.align(readTable, mix4a_keep, bar.ref)
write.csv(mix4a_bar_table, file="../data/Mix4a_barcode_counts.csv", quote=F)

readTable <- MULTIseq.preProcess(R1 = "../fastq/Mix4b-Barcode/R1_paired.fastq.gz",
                                 R2 = "../fastq/Mix4b-Barcode/R2_paired.fastq.gz",
                                 cellIDs = mix4b_keep,
                                 chemistry="V2")
mix4b_bar_table <- MULTIseq.align(readTable, mix4b_keep, bar.ref)
write.csv(mix4b_bar_table, file="../data/Mix4b_barcode_counts.csv", quote=F)
```

## Mix1 Annotation
```{r}
mix1_tsne <- barTSNE(mix1_bar_table[,1:96])
```

```{r}
mix1.temp <- mix1_bar_table
mix1.temp$TSNE1 <- mix1_tsne$TSNE1
mix1.temp$TSNE2 <- mix1_tsne$TSNE2
write.csv(mix1.temp, file="../output/mix1_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix1_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix1_tsne$TSNE1
temp$TSNE2 <- mix1_tsne$TSNE2


for (LMO in colnames(mix1_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix1/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix1_remove <- c(1, 6, 12, 15, 21, 26, 29, 36, 39, 
                 40, 41, 42, 53, 64, 61, 73, 85)
```

```{r}
mix1_bar_filtered <- mix1_bar_table[,1:96]
mix1_bar_filtered <- mix1_bar_filtered[,-mix1_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix1_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```
```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix1_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix1_bar_filtered <- mix1_bar_filtered[-which(rownames(mix1_bar_filtered) %in% neg.cells), ]
```


Note: With this multiplex run, I've tried the multiple rounds of classification, but
the calls after Round 1 are quite good by inspection

Let's quickly look at the TSNE with classifications
```{r}
mix1_tsne$Classification <- "Singlet"
mix1_tsne$Classification[which(round1.calls[rownames(mix1_tsne)]=="Doublet")] <- "Doublet"
mix1_tsne$Classification[which(round1.calls[rownames(mix1_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix1_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix1_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```

The singlets/doublets look fair. We're losing a lot as negatives around fairly obvious clusters.
Let's try to reclassify them

### Negative reclassification
```{r}
mix1_bar_full <- mix1_bar_table[,1:96]
mix1_bar_full <- mix1_bar_full[,-mix1_remove]
reclass.cells <- findReclassCells(mix1_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells_temp(mix1_bar_full, round1.calls[rownames(mix1_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1, ], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 11) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix1_tsne$Classification <- "Singlet"
mix1_tsne$Classification[which(final.calls[rownames(mix1_tsne)]=="Doublet")] <- "Doublet"
mix1_tsne$Classification[which(final.calls[rownames(mix1_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix1_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix1_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, alpha=0.5, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix1_tsne.png",
       width=5.5, height=4)
tsne_classification
```

## Assess sample classifications on TSNE
```{r}
mix1.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix1_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix1_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix1/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix1.samples, plotSampleTSNE)
```

```{r}
mix1_calls <- final.calls
```

## Mix2 Annotation
```{r}
mix2_tsne <- barTSNE(mix2_bar_table[,1:96])
```

```{r}
mix2.temp <- mix2_bar_table
mix2.temp$TSNE1 <- mix2.temp$TSNE1
mix2.temp$TSNE2 <- mix2.temp$TSNE2
write.csv(mix2.temp, file="../output/mix2_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix2_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix2_tsne$TSNE1
temp$TSNE2 <- mix2_tsne$TSNE2


for (LMO in colnames(mix2_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix2/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix2_remove <- c(58, 79, 81,82, 84, 85, 91, 92, 93, 94, 96)
```

```{r}
mix2_bar_filtered <- mix2_bar_table[,1:96]
mix2_bar_filtered <- mix2_bar_filtered[,-mix2_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix2_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```

```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix2_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix2_bar_filtered <- mix2_bar_filtered[-which(rownames(mix2_bar_filtered) %in% neg.cells), ]
```


Note: With this multiplex run, I've tried the multiple rounds of classification, but
the calls after Round 1 are quite good by inspection

Let's quickly look at the TSNE with classifications
```{r}
mix2_tsne$Classification <- "Singlet"
mix2_tsne$Classification[which(round1.calls[rownames(mix2_tsne)]=="Doublet")] <- "Doublet"
mix2_tsne$Classification[which(round1.calls[rownames(mix2_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix2_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix2_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```

The singlets/doublets look fair. We're losing a lot as negatives around fairly obvious clusters.
Let's try to reclassify them

### Negative reclassification
```{r}
mix2_bar_full <- mix2_bar_table[,1:96]
mix2_bar_full <- mix2_bar_full[,-mix2_remove]
reclass.cells <- findReclassCells(mix2_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells_temp(mix2_bar_full, round1.calls[rownames(mix2_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1, ], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 10) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix2_tsne$Classification <- "Singlet"
mix2_tsne$Classification[which(final.calls[rownames(mix2_tsne)]=="Doublet")] <- "Doublet"
mix2_tsne$Classification[which(final.calls[rownames(mix2_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix2_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix2_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix2_tsne.png",
       width=5.5, height=4)
tsne_classification
```


## Assess sample classifications on TSNE
```{r}
mix2.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix2_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix2_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix2/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix2.samples, plotSampleTSNE)
```

```{r}
mix2_calls <- final.calls
```

## Mix3a Annotation
```{r}
mix3a_tsne <- barTSNE(mix3a_bar_table[,1:96])
```

```{r}
mix3a.temp <- mix3a_bar_table
mix3a.temp$TSNE1 <- mix3a_tsne$TSNE1
mix3a.temp$TSNE2 <- mix3a_tsne$TSNE2
write.csv(mix3a.temp, file="../output/mix3a_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix3a_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix3a_tsne$TSNE1
temp$TSNE2 <- mix3a_tsne$TSNE2


for (LMO in colnames(mix3a_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix3a/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix3a_remove <- c(25, 31, 55, 69, 70, 79)
```

```{r}
mix3a_bar_filtered <- mix3a_bar_table[,1:96]
mix3a_bar_filtered <- mix3a_bar_filtered[,-mix3a_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix3a_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```

```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix3a_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix3a_bar_filtered <- mix3a_bar_filtered[-which(rownames(mix3a_bar_filtered) %in% neg.cells), ]
```


Note: With this multiplex run, I've tried the multiple rounds of classification, but
the calls after Round 1 are quite good by inspection

Let's quickly look at the TSNE with classifications
```{r}
mix3a_tsne$Classification <- "Singlet"
mix3a_tsne$Classification[which(round1.calls[rownames(mix3a_tsne)]=="Doublet")] <- "Doublet"
mix3a_tsne$Classification[which(round1.calls[rownames(mix3a_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix3a_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix3a_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```

The singlets/doublets look fair. We're losing a lot as negatives around fairly obvious clusters.
Let's try to reclassify them

### Negative reclassification
```{r}
mix3a_bar_full <- mix3a_bar_table[,1:96]
mix3a_bar_full <- mix3a_bar_full[,-mix3a_remove]
reclass.cells <- findReclassCells(mix3a_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells(mix3a_bar_full, round1.calls[rownames(mix3a_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1, ], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 9) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix3a_tsne$Classification <- "Singlet"
mix3a_tsne$Classification[which(final.calls[rownames(mix3a_tsne)]=="Doublet")] <- "Doublet"
mix3a_tsne$Classification[which(final.calls[rownames(mix3a_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix3a_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix3a_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix3a_tsne.png",
       width=5.5, height=4)
tsne_classification
```

## Assess sample classifications on TSNE
```{r}
mix3a.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix3a_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix3a_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix3a/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix3a.samples, plotSampleTSNE)
```

```{r}
mix3a_calls <- final.calls
```

```{r}
write.csv(mix3a_tsne, file="../output/mix3a_barcode_tsne_classifications.csv",
          quote=F)
```


## Mix3b Annotation
```{r}
mix3b_tsne <- barTSNE(mix3b_bar_table[,1:96])
```

```{r}
mix3b.temp <- mix3b_bar_table
mix3b.temp$TSNE1 <- mix3b_tsne$TSNE1
mix3b.temp$TSNE2 <- mix3b_tsne$TSNE2
write.csv(mix3b.temp, file="../output/mix3b_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix3b_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix3b_tsne$TSNE1
temp$TSNE2 <- mix3b_tsne$TSNE2


for (LMO in colnames(mix3b_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix3b/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix3b_remove <- c(25, 31, 55, 79, 92)
```

```{r}
mix3b_bar_filtered <- mix3b_bar_table[,1:96]
mix3b_bar_filtered <- mix3b_bar_filtered[,-mix3b_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix3b_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```

```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix3b_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix3b_bar_filtered <- mix3b_bar_filtered[-which(rownames(mix3b_bar_filtered) %in% neg.cells), ]
```


Note: With this multiplex run, I've tried the multiple rounds of classification, but
the calls after Round 1 are quite good by inspection

Let's quickly look at the TSNE with classifications
```{r}
mix3b_tsne$Classification <- "Singlet"
mix3b_tsne$Classification[which(round1.calls[rownames(mix3b_tsne)]=="Doublet")] <- "Doublet"
mix3b_tsne$Classification[which(round1.calls[rownames(mix3b_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix3b_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix3b_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```

The singlets/doublets look fair. We're losing a lot as negatives around fairly obvious clusters.
Let's try to reclassify them

### Negative reclassification
```{r}
mix3b_bar_full <- mix3b_bar_table[,1:96]
mix3b_bar_full <- mix3b_bar_full[,-mix3b_remove]
reclass.cells <- findReclassCells(mix3b_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells_temp(mix3b_bar_full, round1.calls[rownames(mix3b_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1, ], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 8) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix3b_tsne$Classification <- "Singlet"
mix3b_tsne$Classification[which(final.calls[rownames(mix3b_tsne)]=="Doublet")] <- "Doublet"
mix3b_tsne$Classification[which(final.calls[rownames(mix3b_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix3b_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix3b_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix3b_tsne.png",
       width=5.5, height=4)
tsne_classification
```

## Assess sample classifications on TSNE
```{r}
mix3b.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix3b_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix3b_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix3b/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix3b.samples, plotSampleTSNE)
```

```{r}
mix3b_calls <- final.calls
```

## Mix4a Annotation
```{r}
mix4a_tsne <- barTSNE(mix4a_bar_table[,1:96])
```

```{r}
mix4a.temp <- mix4a_bar_table
mix4a.temp$TSNE1 <- mix4a_tsne$TSNE1
mix4a.temp$TSNE2 <- mix4a_tsne$TSNE2
write.csv(mix4a.temp, file="../output/mix4a_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix4a_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix4a_tsne$TSNE1
temp$TSNE2 <- mix4a_tsne$TSNE2


for (LMO in colnames(mix4a_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix4a/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix4a_remove <- c(4, 12, 16, 28, 40, 42, 65, 66, 88)
#12 looks like it could be classified but the signal is suuuper low and ends up not
#being classified at all
```

```{r}
mix4a_bar_filtered <- mix4a_bar_table[,1:96]
mix4a_bar_filtered <- mix4a_bar_filtered[,-mix4a_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix4a_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```

```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix4a_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix4a_bar_filtered <- mix4a_bar_filtered[-which(rownames(mix4a_bar_filtered) %in% neg.cells), ]
```

Let's quickly look at the TSNE with classifications
```{r}
mix4a_tsne$Classification <- "Singlet"
mix4a_tsne$Classification[which(round1.calls[rownames(mix4a_tsne)]=="Doublet")] <- "Doublet"
mix4a_tsne$Classification[which(round1.calls[rownames(mix4a_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix4a_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix4a_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```


### Negative reclassification
```{r}
mix4a_bar_full <- mix4a_bar_table[,1:96]
mix4a_bar_full <- mix4a_bar_full[,-mix4a_remove]
reclass.cells <- findReclassCells(mix4a_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells(mix4a_bar_full, round1.calls[rownames(mix4a_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1,], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 7) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix4a_tsne$Classification <- "Singlet"
mix4a_tsne$Classification[which(final.calls[rownames(mix4a_tsne)]=="Doublet")] <- "Doublet"
mix4a_tsne$Classification[which(final.calls[rownames(mix4a_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix4a_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix4a_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix4a_tsne.png",
       width=5.5, height=4)
tsne_classification
```

## Assess sample classifications on TSNE
```{r}
mix4a.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix4a_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix4a_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix4a/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix4a.samples, plotSampleTSNE)
```

```{r}
mix4a_calls <- final.calls
```

## Mix4b Annotation
```{r}
mix4b_tsne <- barTSNE(mix4b_bar_table[,1:96])
```

```{r}
mix4b.temp <- mix4b_bar_table
mix4b.temp$TSNE1 <- mix4b_tsne$TSNE1
mix4b.temp$TSNE2 <- mix4b_tsne$TSNE2
write.csv(mix4b.temp, file="../output/mix4b_barcode_tsne.csv", quote=F)
```

```{r}
temp <- as.matrix(mix4b_tsne[3:98])
temp[temp<0] <- 0
temp <- as.data.frame(temp)
temp$TSNE1 <- mix4b_tsne$TSNE1
temp$TSNE2 <- mix4b_tsne$TSNE2


for (LMO in colnames(mix4b_tsne)[3:98]){
  png(filename = paste0("../figs/LMOs/Mix4b/",LMO,".png"), width = 700, height = 600)
  p <- ggplot(data=temp, aes_string(x="TSNE1",y="TSNE2",color=LMO)) + geom_point() +
    scale_color_gradient(low="lightgrey",high="red") +
    theme_void()
  print(p)
  dev.off()
}
```

Filtering out bad barcodes
```{r}
mix4b_remove <- c(4, 6, 9, 12, 16, 28, 42)
```

```{r}
mix4b_bar_filtered <- mix4b_bar_table[,1:96]
mix4b_bar_filtered <- mix4b_bar_filtered[,-mix4b_remove]
```

```{r}
bar.table_sweep.list <- list()
n <- 0
for (q in seq(0.01, 0.99, by=0.02)) {
  print(q)
  n <- n + 1
  bar.table_sweep.list[[n]] <- classifyCells(mix4b_bar_filtered, q=q)
  names(bar.table_sweep.list)[n] <- paste("q=",q,sep="")
}
```

```{r}
threshold.results1 <- findThresh(call.list=bar.table_sweep.list)
ggplot(data=threshold.results1$res, aes(x=q, y=Proportion, color=Subset)) + geom_line() + theme(legend.position = "none") +
  geom_vline(xintercept=threshold.results1$extrema, lty=2) + scale_color_manual(values=c("red","black","blue"))
```

```{r}
## Finalize round 1 classifications, remove negative cells
round1.calls <- classifyCells(mix4b_bar_filtered, 
                              q=findQ(threshold.results1$res, 
                                      threshold.results1$extrema))
neg.cells <- names(round1.calls)[which(round1.calls == "Negative")]
mix4b_bar_filtered <- mix4b_bar_filtered[-which(rownames(mix4b_bar_filtered) %in% neg.cells), ]
```

Let's quickly look at the TSNE with classifications
```{r}
mix4b_tsne$Classification <- "Singlet"
mix4b_tsne$Classification[which(round1.calls[rownames(mix4b_tsne)]=="Doublet")] <- "Doublet"
mix4b_tsne$Classification[which(round1.calls[rownames(mix4b_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix4b_tsne$Classification)
```


```{r}
tsne_classification <- ggplot(mix4b_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
tsne_classification
```


### Negative reclassification
```{r}
mix4b_bar_full <- mix4b_bar_table[,1:96]
mix4b_bar_full <- mix4b_bar_full[,-mix4b_remove]
reclass.cells <- findReclassCells(mix4b_bar_full,
                                  unique(names(round1.calls)[which(round1.calls=="Negative")]))
reclass.res <- rescueCells(mix4b_bar_full, round1.calls[rownames(mix4b_bar_full)], reclass.cells)
```

```{r}
ggplot(reclass.res[-1,], aes(x=ClassStability, y=MatchRate_mean)) + 
    geom_point() + xlim(c(nrow(reclass.res)-1,1)) + 
    ylim(c(0,1.05)) +
    geom_errorbar(aes(ymin=MatchRate_mean-MatchRate_sd, ymax=MatchRate_mean+MatchRate_sd), width=.1) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1], color="red") +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]+3*reclass.res$MatchRate_sd[1], color="red",lty=2) +
    geom_hline(yintercept = reclass.res$MatchRate_mean[1]-3*reclass.res$MatchRate_sd[1], color="red",lty=2)
```

```{r}
final.calls <- round1.calls
rescue.ind <- which(reclass.cells$ClassStability >= 6) ## Note: Value will be dataset-specific
final.calls[rownames(reclass.cells)[rescue.ind]] <- reclass.cells$Reclassification[rescue.ind]
```

Recheck tSNE
```{r}
mix4b_tsne$Classification <- "Singlet"
mix4b_tsne$Classification[which(final.calls[rownames(mix4b_tsne)]=="Doublet")] <- "Doublet"
mix4b_tsne$Classification[which(final.calls[rownames(mix4b_tsne)]=="Negative")] <- "Negative"
```

```{r}
table(mix4b_tsne$Classification)
```

```{r}
table(final.calls) - table(round1.calls)
```


```{r}
tsne_classification <- ggplot(mix4b_tsne, aes(x=TSNE1, y=TSNE2)) +
  geom_point(size=0.25, aes(color=Classification)) +
  theme_void()
ggsave(tsne_classification, file="../figs/LMOs/Mix4b_tsne.png",
       width=5.5, height=4)
tsne_classification
```

## Assess sample classifications on TSNE
```{r}
mix4b.samples <- unique(final.calls)
plotSampleTSNE <- function(sample){
  data <- mix4b_tsne
  data$Sample <- "Other"
  data$Sample[which(final.calls[rownames(mix4b_tsne)]==sample)] <- sample
  sample_plot <- ggplot(data, aes(x=TSNE1, y=TSNE2)) +
    geom_point(size=0.25, alpha=0.5, aes(color=Sample)) +
    scale_color_manual(values=c("red", "lightgrey")) +
    theme_void()
  ggsave(sample_plot, file=paste0("../figs/LMOs/Mix4b/Classifications/", 
         sample,".png"), width=5, height=3.2)
}
```

```{r}
lapply(mix4b.samples, plotSampleTSNE)
```

```{r}
mix4b_calls <- final.calls
```

# Build annotation table
Goal here is to match the barcode annotations to specific conditions.

This is manually just building the table showing what barcode got added to each well.
Note that we reversed the barcode order between replicates (see Mix1 vs. Mix2 annotations)
```{r}
library(tidyr)
mix2_annotations <- data.frame(barcode=colnames(mix1_bar_table)[1:96])
mix2_annotations$CellLine <- rep(c("A549", "A549", "A549", "DU145", "DU145", "DU145",
                              "MCF7", "MCF7", "MCF7", "OVCA420", "OVCA420", "OVCA420"), 8)
mix2_annotations$Treatment <- rep(c("TGFB1", "EGF", "TNF"), 32)
mix2_annotations$Time <- rep(c("0d", "8h", "1d", "3d", "7d", "8h_rm", "1d_rm", "3d_rm"), each=12)

mix1_annotations <- mix2_annotations
mix1_annotations$CellLine <- rev(mix1_annotations$CellLine)
mix1_annotations$Treatment <- rev(mix1_annotations$Treatment)
mix1_annotations$Time <- rev(mix1_annotations$Time)

mix3a_annotations <- mix2_annotations #Mix 3a/b labelled A1-H12
mix3b_annotations <- mix2_annotations
mix4a_annotations <- mix1_annotations #Mix 4a/b labelled H12-A1
mix4b_annotations <- mix1_annotations
```

```{r}
mix1_annotations$Mix <- "Mix1"
mix1_annotations$CellCount <- table(mix1_calls)[as.character(mix1_annotations$barcode)]
mix1_annotations$CellCount[is.na(mix1_annotations$CellCount)] <- 0

mix2_annotations$Mix <- "Mix2"
mix2_annotations$CellCount <- table(mix2_calls)[as.character(mix2_annotations$barcode)]
mix2_annotations$CellCount[is.na(mix2_annotations$CellCount)] <- 0

mix3a_annotations$Mix <- "Mix3a"
mix3a_annotations$CellCount <- table(mix3a_calls)[as.character(mix3a_annotations$barcode)]
mix3a_annotations$CellCount[is.na(mix3a_annotations$CellCount)] <- 0

mix3b_annotations$Mix <- "Mix3b"
mix3b_annotations$CellCount <- table(mix3b_calls)[as.character(mix3b_annotations$barcode)]
mix3b_annotations$CellCount[is.na(mix3b_annotations$CellCount)] <- 0

mix4a_annotations$Mix <- "Mix4a"
mix4a_annotations$CellCount <- table(mix4a_calls)[as.character(mix4a_annotations$barcode)]
mix4a_annotations$CellCount[is.na(mix4a_annotations$CellCount)] <- 0

mix4b_annotations$Mix <- "Mix4b"
mix4b_annotations$CellCount <- table(mix4b_calls)[as.character(mix4b_annotations$barcode)]
mix4b_annotations$CellCount[is.na(mix4b_annotations$CellCount)] <- 0
```

```{r}
#Master annotation
annotations <- dplyr::bind_rows(mix1_annotations, mix2_annotations, mix3a_annotations,
                                mix3b_annotations, mix4a_annotations, mix4b_annotations)
annotations$Sample <-paste0(annotations$CellLine, "_", annotations$Treatment, "_", annotations$Time)
annotations$Sample <- factor(annotations$Sample, levels=unique(dplyr::arrange(annotations, Sample)$Sample))
```

```{r}
write.csv(annotations, file="../output/annotation_counts.csv", quote=F)
```

## Look at distribution
```{r}
dist.plot <- ggplot(annotations, aes(x=Sample, y=CellCount)) +
  geom_bar(stat="identity", aes(fill=Mix)) +
  geom_hline(yintercept = c(200, 400, 600, 800, 1000, 1200, 1400, 1600, 1800, 2000,
                            2200, 2400, 2600), 
             linetype=2, color="grey") +
  theme_classic() +
  coord_flip()+
  theme(axis.text.x=element_text(angle=45, hjust=1),
        axis.title.y=element_blank())
ggsave(dist.plot, file="../figs/sample_distribution.png", width=4, height=18)
ggsave(dist.plot, file="../figs/sample_distribution.pdf", width=4, height=18)
```

# Map annotations onto Seurat object

```{r}
cell.classifications <- data.frame(cell = c(paste0("Mix1_", names(mix1_calls)),
                                            paste0("Mix2_", names(mix2_calls)),
                                            paste0("Mix3a_", names(mix3a_calls)),
                                            paste0("Mix3b_", names(mix3b_calls)),
                                            paste0("Mix4a_", names(mix4a_calls)),
                                            paste0("Mix4b_", names(mix4b_calls))),
                                   classification = c(paste0("Mix1_",mix1_calls),
                                                      paste0("Mix2_",mix2_calls),
                                                      paste0("Mix3a_",mix3a_calls),
                                                      paste0("Mix3b_",mix3b_calls),
                                                      paste0("Mix4a_",mix4a_calls),
                                                      paste0("Mix4b_",mix4b_calls)))
annotations$Merge <- paste0(annotations$Mix, "_", annotations$barcode)

cell.classifications <- left_join(cell.classifications, annotations,
                                  by=c("classification"="Merge"))
rownames(cell.classifications) <- cell.classifications$cell
```

```{r}
seurat$Sample <- cell.classifications[colnames(seurat),"Sample"]
seurat$CellLine <- cell.classifications[colnames(seurat),"CellLine"]
seurat$Treatment <- cell.classifications[colnames(seurat),"Treatment"]
seurat$Time <- cell.classifications[colnames(seurat),"Time"]

seurat$Doublet <- "Singlet"
seurat$Doublet[which(cell.classifications[colnames(seurat), "classification"] %in% c("Mix1_Doublet", "Mix2_Doublet", "Mix3a_Doublet", "Mix3b_Doublet", "Mix4a_Doublet", "Mix4b_Doublet"))] <- "Doublet"
seurat$Doublet[which(cell.classifications[colnames(seurat), "classification"] %in% c("Mix1_Negative", "Mix2_Negative", "Mix3a_Negative", "Mix3b_Negative", "Mix4a_Negative", "Mix4b_Negative"))] <- "Negative"

seurat$Mix <- Idents(seurat)
```

# Finish basic processing of seurat object
## Normalize
```{r}
seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat, selection.method="vst", nfeatures=3000)
```

## Cell cycle annotation
```{r}
cc.genes <- readLines(con = "~/Data/GeneLists/regev_lab_cell_cycle_genes.txt")
#Split these genes into S markers and G2M markers
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
```

```{r}
seurat <- CellCycleScoring(seurat, s.features = s.genes, 
                           g2m.features = g2m.genes, set.ident = FALSE)
```

## Regress out technical factors
```{r}
seurat <- ScaleData(seurat, vars.to.regress=c("percent.mito", "nCount_RNA",
                                              "S.Score", "G2M.Score"),
                    features=VariableFeatures(seurat))
```

## Clustering and Dimensionality Reduction
```{r}
seurat <- RunPCA(seurat, verbose=F)
seurat <- RunUMAP(seurat, dims = 1:20)
```

```{r}
seurat <- FindNeighbors(seurat, dims=1:20)
seurat <- FindClusters(seurat, resolution=0.05)
```


## Visualization

```{r}
DimPlot(seurat, reduction='umap', group.by="CellLine")
DimPlot(seurat, reduction='umap', group.by="Treatment")
DimPlot(seurat, reduction='umap', group.by="Doublet")
DimPlot(seurat, reduction='umap', group.by="Phase")
DimPlot(seurat, reduction='umap', group.by="Mix")
DimPlot(seurat, reduction='umap')
```



```{r}
saveRDS(seurat, file="../data/seurat_unfiltered.rds")
```




