---
title: "Extended Data Figure 3"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(pheatmap)
library(dplyr)
library(fgsea)
library(Hmisc)
```

# Load the data
```{r}
a549_egf <- readRDS("../data/A549_EGF.rds")
a549_tnf <- readRDS("../data/A549_TNF.rds")
a549_tgfb1 <- readRDS("../data/A549_TGFB1.rds")

du145_egf <- readRDS("../data/DU145_EGF.rds")
du145_tnf <- readRDS("../data/DU145_TNF.rds")
du145_tgfb1 <- readRDS("../data/DU145_TGFB1.rds")

mcf7_egf <- readRDS("../data/MCF7_EGF.rds")
mcf7_tnf <- readRDS("../data/MCF7_TNF.rds")
mcf7_tgfb1 <- readRDS("../data/MCF7_TGFB1.rds")

ovca420_egf <- readRDS("../data/OVCA420_EGF.rds")
ovca420_tnf <- readRDS("../data/OVCA420_TNF.rds")
ovca420_tgfb1 <- readRDS("../data/OVCA420_TGFB1.rds")
```

# Extended Data Figure 3a
## Heatmaps
Load smoothed expression values from GAM model
```{r}
a549_tgfb1_fit <- read.csv("../output/pseudotime/a549_tgfb1_fit.csv", row.names=1)
a549_egf_fit <- read.csv("../output/pseudotime/a549_egf_fit.csv", row.names=1)
a549_tnf_fit <- read.csv("../output/pseudotime/a549_tnf_fit.csv", row.names=1)

du145_tgfb1_fit <- read.csv("../output/pseudotime/du145_tgfb1_fit.csv", row.names=1)
du145_egf_fit <- read.csv("../output/pseudotime/du145_egf_fit.csv", row.names=1)
du145_tnf_fit <- read.csv("../output/pseudotime/du145_tnf_fit.csv", row.names=1)

mcf7_tgfb1_fit <- read.csv("../output/pseudotime/mcf7_tgfb1_fit.csv", row.names=1)
mcf7_egf_fit <- read.csv("../output/pseudotime/mcf7_egf_fit.csv", row.names=1)
mcf7_tnf_fit <- read.csv("../output/pseudotime/mcf7_tnf_fit.csv", row.names=1)

ovca420_tgfb1_fit <- read.csv("../output/pseudotime/ovca420_tgfb1_fit.csv", row.names=1)
ovca420_egf_fit <- read.csv("../output/pseudotime/ovca420_egf_fit.csv", row.names=1)
ovca420_tnf_fit <- read.csv("../output/pseudotime/ovca420_tnf_fit.csv", row.names=1)
```

Load differential expression testing
```{r}
a549_tgfb1_gam <- read.csv("../output/pseudotime/a549_tgfb1_dge.csv", row.names = 1)
a549_egf_gam <- read.csv("../output/pseudotime/a549_egf_dge.csv", row.names=1)
a549_tnf_gam <- read.csv("../output/pseudotime/a549_tnf_dge.csv", row.names=1)

du145_tgfb1_gam <- read.csv("../output/pseudotime/du145_tgfb1_dge.csv", row.names=1)
du145_egf_gam <- read.csv("../output/pseudotime/du145_egf_dge.csv", row.names=1)
du145_tnf_gam <- read.csv("../output/pseudotime/du145_tnf_dge.csv", row.names=1)

mcf7_tgfb1_gam <- read.csv("../output/pseudotime/mcf7_tgfb1_dge.csv", row.names=1)
mcf7_egf_gam <- read.csv("../output/pseudotime/mcf7_egf_dge.csv", row.names=1)
mcf7_tnf_gam <- read.csv("../output/pseudotime/mcf7_tnf_dge.csv", row.names=1)

ovca420_tgfb1_gam <- read.csv("../output/pseudotime/ovca420_tgfb1_dge.csv", row.names=1)
ovca420_egf_gam <- read.csv("../output/pseudotime/ovca420_egf_dge.csv", row.names=1)
ovca420_tnf_gam <- read.csv("../output/pseudotime/ovca420_tnf_dge.csv", row.names=1)
```

```{r}
a549_tgfb1_sig <- rownames(a549_tgfb1_gam)[which(a549_tgfb1_gam$fdr <= 0.05)]
a549_egf_sig <- rownames(a549_egf_gam)[which(a549_egf_gam$fdr <= 0.05)]
a549_tnf_sig <- rownames(a549_tnf_gam)[which(a549_tnf_gam$fdr <= 0.05)]

du145_tgfb1_sig <- rownames(du145_tgfb1_gam)[which(du145_tgfb1_gam$fdr <= 0.05)]
du145_egf_sig <- rownames(du145_egf_gam)[which(du145_egf_gam$fdr <= 0.05)]
du145_tnf_sig <- rownames(du145_tnf_gam)[which(du145_tnf_gam$fdr <= 0.05)]

mcf7_tgfb1_sig <- rownames(mcf7_tgfb1_gam)[which(mcf7_tgfb1_gam$fdr <= 0.05)]
mcf7_egf_sig <- rownames(mcf7_egf_gam)[which(mcf7_egf_gam$fdr <= 0.05)]
mcf7_tnf_sig <- rownames(mcf7_tnf_gam)[which(mcf7_tnf_gam$fdr <= 0.05)]

ovca420_tgfb1_sig <- rownames(ovca420_tgfb1_gam)[which(ovca420_tgfb1_gam$fdr <= 0.05)]
ovca420_egf_sig <- rownames(ovca420_egf_gam)[which(ovca420_egf_gam$fdr <= 0.05)]
ovca420_tnf_sig <- rownames(ovca420_tnf_gam)[which(ovca420_tnf_gam$fdr <= 0.05)]
```

Many genes are significant with very modest effect sizes. To try to make it a little more manageable, we'll filter for only genes that are also in the top 2k variable genes

```{r}
#Make a function so we can tweak the number of genes
getVarGenes <- function(seurat_object, n){
  var_table <- seurat_object[["RNA"]]@meta.features
  var_genes <- rownames(var_table[order(desc(var_table$vst.variance.standardized)),])[1:n]
}
```

```{r}
a549_tgfb1_sig <- a549_tgfb1_sig[which(a549_tgfb1_sig %in% getVarGenes(a549_tgfb1, 2000))]
a549_egf_sig <- a549_egf_sig[which(a549_egf_sig %in% getVarGenes(a549_egf, 2000))]
a549_tnf_sig <- a549_tnf_sig[which(a549_tnf_sig %in% getVarGenes(a549_tnf, 2000))]

a549_tgfb1_sig <- a549_tgfb1_sig[which(a549_tgfb1_sig %in% getVarGenes(a549_tgfb1, 2000))]
a549_egf_sig <- a549_egf_sig[which(a549_egf_sig %in% getVarGenes(a549_egf, 2000))]
a549_tnf_sig <- a549_tnf_sig[which(a549_tnf_sig %in% getVarGenes(a549_tnf, 2000))]

du145_tgfb1_sig <- du145_tgfb1_sig[which(du145_tgfb1_sig %in% getVarGenes(du145_tgfb1, 2000))]
du145_egf_sig <- du145_egf_sig[which(du145_egf_sig %in% getVarGenes(du145_egf, 2000))]
du145_tnf_sig <- du145_tnf_sig[which(du145_tnf_sig %in% getVarGenes(du145_tnf, 2000))]

mcf7_tgfb1_sig <- mcf7_tgfb1_sig[which(mcf7_tgfb1_sig %in% getVarGenes(mcf7_tgfb1, 2000))]
mcf7_egf_sig <- mcf7_egf_sig[which(mcf7_egf_sig %in% getVarGenes(mcf7_egf, 2000))]
mcf7_tnf_sig <- mcf7_tnf_sig[which(mcf7_tnf_sig %in% getVarGenes(mcf7_tnf, 2000))]

ovca420_tgfb1_sig <- ovca420_tgfb1_sig[which(ovca420_tgfb1_sig %in% getVarGenes(ovca420_tgfb1, 2000))]
ovca420_egf_sig <- ovca420_egf_sig[which(ovca420_egf_sig %in% getVarGenes(ovca420_egf, 2000))]
ovca420_tnf_sig <- ovca420_tnf_sig[which(ovca420_tnf_sig %in% getVarGenes(ovca420_tnf, 2000))]
```

Make heatmap
```{r}
plotHeatmap <- function(smooth_fit, genes, file_dir){
  exp <- smooth_fit[genes,]
  exp[exp>0.5] <- 0.5
  exp[exp<(-0.5)] <- (-0.5)
  print("Making heatmap")
  heatmap <- pheatmap(exp,
                      color=viridis::magma(100),
                      cluster_cols = F,
                      cluster_rows = T,
                      treeheight_row = 0,
                      clustering_method="ward.D2",
                      legend=F,
                      show_rownames=F,
                      show_colnames=F,
                      filename=paste0("../figure_panels/ExtendedDataFig_3/", file_dir, 
                                      "_Pseudotime_Heatmap.png"),
                      height=6, width=3)
}
```

```{r}
du145_tgfb1_heatmap <- plotHeatmap(du145_tgfb1_fit, du145_tgfb1_sig, "DU145_TGFB1")
du145_egf_heatmap <- plotHeatmap(du145_egf_fit, du145_egf_sig, "DU145_EGF")
du145_tnf_heatmap <- plotHeatmap(du145_tnf_fit, du145_tnf_sig, "DU145_TNF")

mcf7_tgfb1_heatmap <- plotHeatmap(mcf7_tgfb1_fit, mcf7_tgfb1_sig, "MCF7_TGFB1")
mcf7_egf_heatmap <- plotHeatmap(mcf7_egf_fit, mcf7_egf_sig, "MCF7_EGF")
mcf7_tnf_heatmap <- plotHeatmap(mcf7_tnf_fit, mcf7_tnf_sig, "MCF7_TNF")

ovca420_tgfb1_heatmap <- plotHeatmap(ovca420_tgfb1_fit, ovca420_tgfb1_sig, "OVCA420_TGFB1")
ovca420_egf_heatmap <- plotHeatmap(ovca420_egf_fit, ovca420_egf_sig, "OVCA420_EGF")
ovca420_tnf_heatmap <- plotHeatmap(ovca420_tnf_fit, ovca420_tnf_sig, "OVCA420_TNF")
```

# Extended Data Figure 3b
## GSEA plots of DEGs

```{r}
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
```

```{r}
rename_pathways <- function(x){
  s = unlist(strsplit(x, split="_", fixed=T))[-1] #gets rid of first part of name and underscores
  s = Hmisc::capitalize(tolower(s)) #capitalizes first letter of each word, rest are lower case
  s = paste(s, collapse=" ") #bring words together with space between
  return(s)
}

names(hallmarks) <- unlist(lapply(names(hallmarks), rename_pathways))
```

```{r}
plotEMT <- function(sig_list, file_dir){
  sig_list <- arrange(sig_list, desc(fdr))
  rank <- 1:nrow(sig_list)
  names(rank) <- sig_list$Gene
  gsea_plot <- fgsea::plotEnrichment(hallmarks[["Epithelial Mesenchymal Transition"]], rank) +
    #labs(title="Epithelial Mesenchymal Transition") + 
    ylab("ES") + xlab("Gene Variance Rank") +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.line = element_line(size=0.5),
          axis.ticks = element_blank())
  ggsave(gsea_plot, filename = paste0("../figure_panels/ExtendedDataFig_3/", file_dir, "_DEG_GSEA.pdf"),
         width=1.75, height=1.5)
  gsea_plot
}
```

```{r}
plotEMT(a549_tgfb1_gam, "A549_TGFB1")
plotEMT(a549_egf_gam, "A549_EGF")
plotEMT(a549_tnf_gam, "A549_TNF")

plotEMT(du145_tgfb1_gam, "DU145_TGFB1")
plotEMT(du145_egf_gam, "DU145_EGF")
plotEMT(du145_tnf_gam, "DU145_TNF")

plotEMT(mcf7_tgfb1_gam, "MCF7_TGFB1")
plotEMT(mcf7_egf_gam, "MCF7_EGF")
plotEMT(mcf7_tnf_gam, "MCF7_TNF")

plotEMT(ovca420_tgfb1_gam, "OVCA420_TGFB1")
plotEMT(ovca420_egf_gam, "OVCA420_EGF")
plotEMT(ovca420_tnf_gam, "OVCA420_TNF")
```

# Extended Data Figure 3
## Heatmaps of EMT hallmark genes
```{r}
smoothValues <- function(seurat_object, genes, file_dir){
  genes_keep <- genes[which(genes %in% rownames(seurat_object))]
  print(paste0("Genes not found in the data: ", genes[which(genes %nin% rownames(seurat_object))]))
  exp <- as.matrix(seurat_object[["RNA"]]@data[genes_keep,])
  
  fits <- apply(exp, 1, function(z){
    pseudo_data <- data.frame(Gene=z,
                              Pseudotime=seurat_object$Pseudotime)
    model <- mgcv::gam(Gene ~ s(Pseudotime, k=4), data=pseudo_data,
                       method="REML")
    target_values <- data.frame(Pseudotime = seq(0, 1, length.out=201))
    new_pseudo_values <- predict(model, newdata=target_values)
  })
  fits <- t(fits)
}

smoothHeatmap <- function(seurat_object, genes, file_dir, height){
  print("Modelling smooth values...")
  mat <- smoothValues(seurat_object, genes)
  print("Scaling...")
  mat <- t(scale(t(mat), scale=F, center=T))
  
  #Add value of zero for absent genes
  #Though this only works if zero is meaningful :/ 
  #Maybe can set it to -2?
  #genes_absent <- genes[which(genes %nin% rownames(seurat_object))]
  #empty_mat <- matrix(0, nrow=length(genes_absent), ncol=ncol(mat))
  #rownames(empty_mat) <- genes_absent
  
  print("Making heatmap...")
  mat[mat>0.5] <- 0.5
  mat[mat<(-0.5)] <- (-0.5)
  
  #bluered <- colorRampPalette(rev(RColorBrewer::brewer.pal(n = 11, "RdBu")))(100)
  heatmap <- pheatmap(mat,
                      color=viridis::inferno(100),
                      #color=bluered,
                      cluster_cols=F,
                      cluster_rows=T,
                      show_colnames=F,
                      show_rownames=T,
                      border_color=NA,
                      legend=F,
                      clustering_method="ward.D2",
                                  filename=paste0("../figure_panels/ExtendedDataFig_3/",file_dir,"_emt_hallmarks_heatmap.png"),
                      width=6, height=19)
}
```

```{r}
smoothHeatmap(a549_egf, hallmarks[["Epithelial Mesenchymal Transition"]], "A549_EGF")
smoothHeatmap(ovca420_tnf, hallmarks[["Epithelial Mesenchymal Transition"]], "OVCA420_TNF")
```




