---
title: "Figure 2"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(cowplot)
library(ggplot2)
library(dplyr)
library(viridis)
library(ggforce)
library(pheatmap)
library(fgsea)
library(ggsci)
library(RColorBrewer)
```

# Load the data

```{r}
a549_egf <- readRDS("../data/A549_EGF.rds")
a549_tnf <- readRDS("../data/A549_TNF.rds")
a549_tgfb1 <- readRDS("../data/A549_TGFB1.rds")

du145_egf <- readRDS("../data/DU145_EGF.rds")
du145_tnf <- readRDS("../data/DU145_TNF.rds")
du145_tgfb1 <- readRDS("../data/DU145_TGFB1.rds")

mcf7_egf <- readRDS("../data/MCF7_EGF.rds")
mcf7_tnf <- readRDS("../data/MCF7_TNF.rds")
mcf7_tgfb1 <- readRDS("../data/MCF7_TGFB1.rds")

ovca420_egf <- readRDS("../data/OVCA420_EGF.rds")
ovca420_tnf <- readRDS("../data/OVCA420_TNF.rds")
ovca420_tgfb1 <- readRDS("../data/OVCA420_TGFB1.rds")
```

# Figure 2a
```{r}
plot_data <- a549_tgfb1@meta.data
plot_data$UMAP1 <- Embeddings(a549_tgfb1, "umap_pseudo")[,1]
plot_data$UMAP2 <- Embeddings(a549_tgfb1, "umap_pseudo")[,2]
```

```{r}
PRGn <- rev(RColorBrewer::brewer.pal(11, "PRGn"))
color_ramp <- c(PRGn[1], PRGn[2], PRGn[3], PRGn[4], "grey50", PRGn[7], PRGn[8], PRGn[9])
```


```{r}
timepoint_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.25, alpha=0.75, aes(color=Sample),
             shape=16) +
  scale_color_manual(values=color_ramp) +
  theme_void() +
  theme(legend.position="none")
ggsave(timepoint_plot, filename = "../figure_panels/figure_2/umap_timepoint.png",
       width=2.75, height=1.75, dpi=600)

pseudo_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.25, alpha=0.75, aes(color=Pseudotime),
             shape=16) +
  scale_color_viridis(option="D") +
  theme_void() +
  theme(legend.title=element_blank(),
        legend.text=element_blank())
ggsave(pseudo_plot, filename = "../figure_panels/figure_2/umap_pseudotime_withLegend.png",
       width=2.75, height=1.75, dpi=600)

pseudo_plot <- ggplot(plot_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.25, alpha=0.75, aes(color=Pseudotime),
             shape=16) +
  scale_color_viridis(option="D") +
  theme_void() +
  theme(legend.position="none")
ggsave(pseudo_plot, filename = "../figure_panels/figure_2/umap_pseudotime_noLegend.png",
       width=2.75, height=1.75, dpi=600)
```

```{r}
timepoint_plot
pseudo_plot
```

# Figure 2b
## Sina plots of pseudotime
```{r}
plotPseudotime <- function(seurat_object, file_dir){
  data <- seurat_object@meta.data
  data_plot <- ggplot(data, aes(x=Sample, y=Pseudotime)) +
    geom_sina(size=0.01, alpha=0.75, aes(color=Sample), shape=16) +
    stat_summary(fun.y=median, fun.ymin=median, fun.ymax=median,
                 geom='crossbar', width=0.5, fatten=0, size=1) +
    scale_color_manual(values=color_ramp, labels=c("0d", "8h", "1d", "3d", "7d",
                                                   "7d + 8h off", "7d + 1d off",
                                                   "7d + 3d off")) +
    theme_classic() +
    theme(legend.position="none",
          axis.text=element_blank(),
          axis.title=element_blank(),
          axis.line = element_line(size=0.3),
          axis.ticks = element_line(size=0.3))
}
```

```{r}
p1 <- plotPseudotime(a549_tgfb1, "A549_TGFB1")
p2 <- plotPseudotime(a549_egf, "A549_EGF")
p3 <- plotPseudotime(a549_tnf, "A549_TNF")

p4 <- plotPseudotime(du145_tgfb1, "DU145_TGFB1")
p5 <- plotPseudotime(du145_egf, "DU145_EGF")
p6 <- plotPseudotime(du145_tnf, "DU145_TNF")

p7 <- plotPseudotime(mcf7_tgfb1, "MCF7_TGFB1")
p8 <- plotPseudotime(mcf7_egf, "MCF7_EGF")
p9 <- plotPseudotime(mcf7_tnf, "MCF7_TNF")

p10 <- plotPseudotime(ovca420_tgfb1, "OVCA420_TGFB1")
p11 <- plotPseudotime(ovca420_egf, "OVCA420_EGF")
p12 <- plotPseudotime(ovca420_tnf, "OVCA420_TNF")
```

```{r}
grid_plots <- cowplot::plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12,
                        ncol=3)
cowplot::save_plot(grid_plots, filename="../figure_panels/figure_2/Pseudotime_Sina.png",
                   base_width=15, base_height=12, dpi=600, units="cm")
```


# Figure 2c
## Three heatmaps
Load smoothed expression values from GAM model
```{r}
a549_tgfb1_fit <- read.csv("../output/pseudotime/a549_tgfb1_fit.csv", row.names=1)
a549_egf_fit <- read.csv("../output/pseudotime/a549_egf_fit.csv", row.names=1)
a549_tnf_fit <- read.csv("../output/pseudotime/a549_tnf_fit.csv", row.names=1)
```

Load differential expression testing
```{r}
a549_tgfb1_gam <- read.csv("../output/pseudotime/a549_tgfb1_dge.csv", row.names = 1)
a549_egf_gam <- read.csv("../output/pseudotime/a549_egf_dge.csv", row.names=1)
a549_tnf_gam <- read.csv("../output/pseudotime/a549_tnf_dge.csv", row.names=1)

du145_tgfb1_gam <- read.csv("../output/pseudotime/du145_tgfb1_dge.csv", row.names=1)
du145_egf_gam <- read.csv("../output/pseudotime/du145_egf_dge.csv", row.names=1)
du145_tnf_gam <- read.csv("../output/pseudotime/du145_tnf_dge.csv", row.names=1)

mcf7_tgfb1_gam <- read.csv("../output/pseudotime/mcf7_tgfb1_dge.csv", row.names=1)
mcf7_egf_gam <- read.csv("../output/pseudotime/mcf7_egf_dge.csv", row.names=1)
mcf7_tnf_gam <- read.csv("../output/pseudotime/mcf7_tnf_dge.csv", row.names=1)

ovca420_tgfb1_gam <- read.csv("../output/pseudotime/ovca420_tgfb1_dge.csv", row.names=1)
ovca420_egf_gam <- read.csv("../output/pseudotime/ovca420_egf_dge.csv", row.names=1)
ovca420_tnf_gam <- read.csv("../output/pseudotime/ovca420_tnf_dge.csv", row.names=1)
```

```{r}
a549_tgfb1_sig <- rownames(a549_tgfb1_gam)[which(a549_tgfb1_gam$fdr <= 0.05)]
a549_egf_sig <- rownames(a549_egf_gam)[which(a549_egf_gam$fdr <= 0.05)]
a549_tnf_sig <- rownames(a549_tnf_gam)[which(a549_tnf_gam$fdr <= 0.05)]

du145_tgfb1_sig <- rownames(du145_tgfb1_gam)[which(du145_tgfb1_gam$fdr <= 0.05)]
du145_egf_sig <- rownames(du145_egf_gam)[which(du145_egf_gam$fdr <= 0.05)]
du145_tnf_sig <- rownames(du145_tnf_gam)[which(du145_tnf_gam$fdr <= 0.05)]

mcf7_tgfb1_sig <- rownames(mcf7_tgfb1_gam)[which(mcf7_tgfb1_gam$fdr <= 0.05)]
mcf7_egf_sig <- rownames(mcf7_egf_gam)[which(mcf7_egf_gam$fdr <= 0.05)]
mcf7_tnf_sig <- rownames(mcf7_tnf_gam)[which(mcf7_tnf_gam$fdr <= 0.05)]

ovca420_tgfb1_sig <- rownames(ovca420_tgfb1_gam)[which(ovca420_tgfb1_gam$fdr <= 0.05)]
ovca420_egf_sig <- rownames(ovca420_egf_gam)[which(ovca420_egf_gam$fdr <= 0.05)]
ovca420_tnf_sig <- rownames(ovca420_tnf_gam)[which(ovca420_tnf_gam$fdr <= 0.05)]
```

Many genes are significant with very modest effect sizes. To try to make it a little more manageable, we'll filter for only genes that are also in the top 2k variable genes

```{r}
#Make a function so we can tweak the number of genes
getVarGenes <- function(seurat_object, n){
  var_table <- seurat_object[["RNA"]]@meta.features
  var_genes <- rownames(var_table[order(desc(var_table$vst.variance.standardized)),])[1:n]
}
```

```{r}
a549_tgfb1_sig <- a549_tgfb1_sig[which(a549_tgfb1_sig %in% getVarGenes(a549_tgfb1, 2000))]
a549_egf_sig <- a549_egf_sig[which(a549_egf_sig %in% getVarGenes(a549_egf, 2000))]
a549_tnf_sig <- a549_tnf_sig[which(a549_tnf_sig %in% getVarGenes(a549_tnf, 2000))]

a549_tgfb1_sig <- a549_tgfb1_sig[which(a549_tgfb1_sig %in% getVarGenes(a549_tgfb1, 2000))]
a549_egf_sig <- a549_egf_sig[which(a549_egf_sig %in% getVarGenes(a549_egf, 2000))]
a549_tnf_sig <- a549_tnf_sig[which(a549_tnf_sig %in% getVarGenes(a549_tnf, 2000))]

du145_tgfb1_sig <- du145_tgfb1_sig[which(du145_tgfb1_sig %in% getVarGenes(du145_tgfb1, 2000))]
du145_egf_sig <- du145_egf_sig[which(du145_egf_sig %in% getVarGenes(du145_egf, 2000))]
du145_tnf_sig <- du145_tnf_sig[which(du145_tnf_sig %in% getVarGenes(du145_tnf, 2000))]

mcf7_tgfb1_sig <- mcf7_tgfb1_sig[which(mcf7_tgfb1_sig %in% getVarGenes(mcf7_tgfb1, 2000))]
mcf7_egf_sig <- mcf7_egf_sig[which(mcf7_egf_sig %in% getVarGenes(mcf7_egf, 2000))]
mcf7_tnf_sig <- mcf7_tnf_sig[which(mcf7_tnf_sig %in% getVarGenes(mcf7_tnf, 2000))]

ovca420_tgfb1_sig <- ovca420_tgfb1_sig[which(ovca420_tgfb1_sig %in% getVarGenes(ovca420_tgfb1, 2000))]
ovca420_egf_sig <- ovca420_egf_sig[which(ovca420_egf_sig %in% getVarGenes(ovca420_egf, 2000))]
ovca420_tnf_sig <- ovca420_tnf_sig[which(ovca420_tnf_sig %in% getVarGenes(ovca420_tnf, 2000))]
```

Make heatmap
```{r}
plotHeatmap <- function(smooth_fit, genes, file_dir){
  exp <- smooth_fit[genes,]
  exp[exp>0.5] <- 0.5
  exp[exp<(-0.5)] <- (-0.5)
  print("Making heatmap")
  heatmap <- pheatmap(exp,
                      color=viridis::magma(100),
                      cluster_cols = F,
                      cluster_rows = T,
                      treeheight_row = 0,
                      clustering_method="ward.D2",
                      legend=F,
                      show_rownames=F,
                      show_colnames=F,
                      filename=paste0("../figure_panels/figure_2/", file_dir, 
                                      "_Pseudotime_Heatmap.png"),
                      height=6, width=3)
}
```

```{r}
a549_tgfb1_heatmap <- plotHeatmap(a549_tgfb1_fit, a549_tgfb1_sig, "A549_TGFB1")
a549_egf_heatmap <- plotHeatmap(a549_egf_fit, a549_egf_sig, "A549_EGF")
a549_tnf_heatmap <- plotHeatmap(a549_tnf_fit, a549_tnf_sig, "A549_TNF")
```


# Figure 2d
## EMT Score Over Pseudotime
```{r}
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
```

```{r}
rename_pathways <- function(x){
  s = unlist(strsplit(x, split="_", fixed=T))[-1] #gets rid of first part of name and underscores
  s = Hmisc::capitalize(tolower(s)) #capitalizes first letter of each word, rest are lower case
  s = paste(s, collapse=" ") #bring words together with space between
  return(s)
}

names(hallmarks) <- unlist(lapply(names(hallmarks), rename_pathways))
```

```{r}
scoreEMT <- function(seurat_object){
  seurat_object <- AddModuleScore(seurat_object, 
                                  features = hallmarks["Epithelial Mesenchymal Transition"],
                                  name="EMT_Score")
  return(seurat_object)
}
```

```{r}
a549_tgfb1 <- scoreEMT(a549_tgfb1)
a549_egf <- scoreEMT(a549_egf)
a549_tnf <- scoreEMT(a549_tnf)

du145_tgfb1 <- scoreEMT(du145_tgfb1)
du145_egf <- scoreEMT(du145_egf)
du145_tnf <- scoreEMT(du145_tnf)

mcf7_tgfb1 <- scoreEMT(mcf7_tgfb1)
mcf7_egf <- scoreEMT(mcf7_egf)
mcf7_tnf <- scoreEMT(mcf7_tnf)

ovca420_tgfb1 <- scoreEMT(ovca420_tgfb1)
ovca420_egf <- scoreEMT(ovca420_egf)
ovca420_tnf <- scoreEMT(ovca420_tnf)
```

```{r}
plotGeneSetScore <- function(geneset){
  data <- bind_rows(a549_tgfb1@meta.data, a549_egf@meta.data, a549_tnf@meta.data,
                    du145_tgfb1@meta.data, du145_egf@meta.data, du145_tnf@meta.data,
                    mcf7_tgfb1@meta.data, mcf7_egf@meta.data, mcf7_tnf@meta.data,
                    ovca420_tgfb1@meta.data, ovca420_egf@meta.data, ovca420_tnf@meta.data)
  #data$Group <- paste0(data$CellLine, "_", data$Treatment)
  shuffle <- sample(1:nrow(data), replace=F)
  data <- data[shuffle,]
  data$Treatment <- factor(data$Treatment, levels=c("TGFB1", "EGF", "TNF"))
  
  data_plot <- ggplot(data, aes_string(x="Pseudotime", y=paste0(geneset, "1"))) +
    #geom_point(size=0.1, alpha=0.25, color="grey20") +
    geom_smooth(size=0.5, aes(color=Treatment)) +
    scale_x_continuous(expand=c(0,0), breaks = c(0,0.5,1)) +
    scale_y_continuous(expand=c(0,0)) +
    scale_color_nejm() +
    ylab("EMT Score") +
    facet_wrap(~CellLine, scales="free", ncol = 2) +
    theme_classic() +
    theme(strip.background = element_blank(),
          strip.text = element_blank(), #I want to put the labels in myself
          legend.position="none",
          axis.title=element_blank(),
          axis.text = element_text(size=6, color="black"),
          axis.line = element_line(size=0.15),
          axis.ticks = element_line(size=0.15))
  data_plot
}
```

```{r}
emt_plot <- plotGeneSetScore("EMT_Score")
emt_plot
```

```{r}
ggsave(emt_plot, filename = "../figure_panels/figure_2/EMT_Score.png",
       units="cm", width=4.5, height=5.2, dpi=600)
```


# Figure 2e
## Sig gene counts
```{r}
deg_list <- data.frame(Gene = unique(c(a549_tgfb1_sig , a549_egf_sig, a549_tnf_sig,
                                       du145_tgfb1_sig, du145_egf_sig, du145_tnf_sig,
                                       mcf7_tgfb1_sig, mcf7_egf_sig, mcf7_tnf_sig,
                                       ovca420_tgfb1_sig, ovca420_egf_sig, ovca420_tnf_sig)))
  
deg_list$A549_TGFB1 <- deg_list$Gene %in% a549_tgfb1_sig
deg_list$A549_EGF <- deg_list$Gene %in% a549_egf_sig
deg_list$A549_TNF <- deg_list$Gene %in% a549_tnf_sig

deg_list$DU145_TGFB1 <- deg_list$Gene %in% du145_tgfb1_sig
deg_list$DU145_EGF <- deg_list$Gene %in% du145_egf_sig
deg_list$DU145_TNF <- deg_list$Gene %in% du145_tnf_sig

deg_list$MCF7_TGFB1 <- deg_list$Gene %in% mcf7_tgfb1_sig
deg_list$MCF7_EGF <- deg_list$Gene %in% mcf7_egf_sig
deg_list$MCF7_TNF <- deg_list$Gene %in% mcf7_tnf_sig

deg_list$OVCA420_TGFB1 <- deg_list$Gene %in% ovca420_tgfb1_sig
deg_list$OVCA420_EGF <- deg_list$Gene %in% ovca420_egf_sig
deg_list$OVCA420_TNF <- deg_list$Gene %in% ovca420_tnf_sig

#Convert TRUE/FALSE to 1/0
deg_list[,2:ncol(deg_list)] <- lapply(deg_list[,2:ncol(deg_list)], as.numeric)
```

```{r}
deg_list_counts <- as.data.frame(rowSums(deg_list[2:ncol(deg_list)]))
colnames(deg_list_counts) <- "Count"
deg_list_counts$Gene <- deg_list$Gene
deg_list_counts <- arrange(deg_list_counts, desc(Count))
deg_list_counts$Index <- 1:nrow(deg_list_counts)
```

```{r}
count_plot <- ggplot(deg_list_counts, aes(x=Index, y=Count)) +
  geom_point(size=0.25, color="black") +
  xlab("") + ylab("Significant Counts") +
  scale_y_continuous(breaks=c(0, 3, 6, 9, 12)) +
  theme_classic() +
  theme(axis.text=element_text(size=5, color="black"),
        axis.title=element_blank(),
        axis.line = element_line(size=0.25),
        axis.ticks = element_line(size=0.25))
ggsave(count_plot, filename = "../figure_panels/figure_2/Diff_Gene_Counts.png",
       dpi=600, width=3.2, height=3.5, units="cm")
```

```{r}
count_plot
```

```{r}
deg_list_counts
```


# Figure 2f
## Heatmap of gene set
```{r}
conserved_genes <- as.character(filter(deg_list_counts, Count >= 6)$Gene)
length(conserved_genes)
```

```{r}
linearGAM <- function(seurat_object, genes){
  genes <- genes[which(genes %in% rownames(seurat_object))]
  exp <- as.matrix(seurat_object[["RNA"]]@data[genes,])
  
  gam.coef <- apply(exp,1,function(z){
    d <- data.frame(exp=z, pseudotime=seurat_object$Pseudotime,
                    batch = seurat_object$Mix)
    tmp <- mgcv::gam(exp ~ pseudotime + batch, data=d)
    p <- coef(tmp)[2] # where the pseudotime coefficient is stored
    p
    })
  gam.coef <- as.data.frame(gam.coef)
  colnames(gam.coef) <- "Coefficient"
  return(gam.coef)
}
```

```{r}
a549_tgfb1_conserved <- linearGAM(a549_tgfb1, conserved_genes)
a549_egf_conserved <- linearGAM(a549_egf, conserved_genes)
a549_tnf_conserved <- linearGAM(a549_tnf, conserved_genes)

du145_tgfb1_conserved <- linearGAM(du145_tgfb1, conserved_genes)
du145_egf_conserved <- linearGAM(du145_egf, conserved_genes)
du145_tnf_conserved <- linearGAM(du145_tnf, conserved_genes)

mcf7_tgfb1_conserved <- linearGAM(mcf7_tgfb1, conserved_genes)
mcf7_egf_conserved <- linearGAM(mcf7_egf, conserved_genes)
mcf7_tnf_conserved <- linearGAM(mcf7_tnf, conserved_genes)

ovca420_tgfb1_conserved <- linearGAM(ovca420_tgfb1, conserved_genes)
ovca420_egf_conserved <- linearGAM(ovca420_egf, conserved_genes)
ovca420_tnf_conserved <- linearGAM(ovca420_tnf, conserved_genes)
```

```{r}
coef_mat <- data.frame(A549_TGFB1 = a549_tgfb1_conserved[conserved_genes,"Coefficient"],
                       A549_EGF = a549_egf_conserved[conserved_genes, "Coefficient"],
                       A549_TNF = a549_tnf_conserved[conserved_genes, "Coefficient"],
                       DU145_TGFB1 = du145_tgfb1_conserved[conserved_genes,"Coefficient"],
                       DU145_EGF = du145_egf_conserved[conserved_genes, "Coefficient"],
                       DU145_TNF = du145_tnf_conserved[conserved_genes, "Coefficient"],
                       MCF7_TGFB1 = mcf7_tgfb1_conserved[conserved_genes,"Coefficient"],
                       MCF7_EGF = mcf7_egf_conserved[conserved_genes, "Coefficient"],
                       MCF7_TNF = mcf7_tnf_conserved[conserved_genes, "Coefficient"],
                       OVCA420_TGFB1 = ovca420_tgfb1_conserved[conserved_genes,"Coefficient"],
                       OVCA420_EGF = ovca420_egf_conserved[conserved_genes, "Coefficient"],
                       OVCA420_TNF = ovca420_tnf_conserved[conserved_genes, "Coefficient"])
coef_mat <- as.matrix(coef_mat)
rownames(coef_mat) <- conserved_genes
coef_mat[is.na(coef_mat)] <- 0

#Only keep values for significant changes
deg_list_temp <- as.matrix(deg_list[,2:ncol(deg_list)])
rownames(deg_list_temp) <- deg_list$Gene
deg_list_temp <- deg_list_temp[rownames(coef_mat),]

sig_mat <- coef_mat * deg_list_temp

# Only keep genes differentially expressed in the same direction in >=6 samples
sig_mat_up <- sig_mat
sig_mat_up[sig_mat_up>0] <- 1
sig_mat_up[sig_mat_up<0] <- 0 
up_genes <- rownames(sig_mat_up)[which(rowSums(sig_mat_up) >= 8)]

sig_mat_down <- sig_mat
sig_mat_down[sig_mat_down>0] <- 0
sig_mat_down[sig_mat_down<0] <- 1 
down_genes <- rownames(sig_mat_down)[which(rowSums(sig_mat_down) >= 8)]

coef_mat <- coef_mat[c(up_genes, down_genes),]
coef_mat[coef_mat>0.25] <- 0.25
coef_mat[coef_mat< (-0.25)] <- (-0.25)
```

```{r}
coef_heatmap <- pheatmap(t(coef_mat),
                         color = colorRampPalette(rev(brewer.pal(7, "RdBu")))(100),
                         border_color = "black",
                         cluster_cols=T,
                         cluster_rows=F,
                         treeheight_col = 0,
                         clustering_method = "ward.D2",
                         show_rownames=F,
                         show_colnames=T,
                         legend=T,
                         cutree_cols=2,
                         filename="../figure_panels/figure_2/blue_red_colorkey.png",
                         width=16, height=4)
```


```{r}
write.csv(up_genes, file="../output/conserved_upregulated_genes.csv", quote=F)
write.csv(down_genes, file="../output/conserved_downregulated_genes.csv", quote=F)
```

# Session Info
```{r}
sessionInfo()
```

